\documentclass[11pt,letterpaper]{article}

% ── Geometry ──────────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}

% ── Pagination ────────────────────────────────────────────────────────────────
\usepackage[all]{nowidow}    % Require ≥2 lines on each side of a page break
\brokenpenalty=10000         % No hyphenated words across page breaks
\raggedbottom                % Allow uneven page bottoms rather than stretching

% ── Fonts ─────────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage{newpxtext}     % Palatino-based serif (TeX Gyre Pagella)
\usepackage{newpxmath}     % Matching math font

% ── Colors ────────────────────────────────────────────────────────────────────
\usepackage[dvipsnames]{xcolor}
\definecolor{SectionBlue}{HTML}{1B3A5C}
\definecolor{AccentGray}{HTML}{4A4A4A}
\definecolor{QuoteBorder}{HTML}{8B9DAF}
\definecolor{RiskRed}{HTML}{C0392B}
\definecolor{RiskAmber}{HTML}{E67E22}
\definecolor{RiskGreen}{HTML}{27AE60}
\definecolor{BoxBg}{HTML}{F7F8FA}

% ── Tables ────────────────────────────────────────────────────────────────────
\usepackage{booktabs}
\usepackage{tabularx}

% ── Boxes ─────────────────────────────────────────────────────────────────────
\usepackage[framemethod=tikz]{mdframed}

% ── Lists ─────────────────────────────────────────────────────────────────────
\usepackage{enumitem}
\usepackage{needspace}

% ── Hyperlinks ────────────────────────────────────────────────────────────────
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=SectionBlue,
  urlcolor=SectionBlue,
  citecolor=SectionBlue,
  pdfauthor={Jim Freeman},
  pdftitle={PR/FAQ: ethos},
  pdfsubject={Working Backwards --- Identity and Workflow for Human-Agent Teams},
}

% ── Bibliography ─────────────────────────────────────────────────────────────
\usepackage[style=numeric,backend=biber,sorting=none]{biblatex}
\addbibresource{prfaq.bib}

% ── Section styling ───────────────────────────────────────────────────────────
\usepackage{titlesec}
\titleformat{\section}
  {\needspace{6\baselineskip}\Large\bfseries\color{SectionBlue}}
  {}
  {0pt}
  {}

\titleformat{\subsection}
  {\needspace{5\baselineskip}\large\bfseries\color{AccentGray}}
  {}
  {0pt}
  {}

\titleformat{\subsubsection}
  {\normalsize\bfseries\color{AccentGray}}
  {}
  {0pt}
  {}
\titlespacing*{\subsubsection}{0pt}{1em}{0.5em}[0pt]

% ── Document stage ───────────────────────────────────────────────────────────
\newcommand{\prfaqstage}[1]{\def\theprfaqstage{#1}}
\prfaqstage{hypothesis}

% ── Document version ─────────────────────────────────────────────────────────
\newcommand{\prfaqversion}[2]{\def\theprfaqmajor{#1}\def\theprfaqminor{#2}}
\prfaqversion{2}{4}

% ── Header/footer ────────────────────────────────────────────────────────────
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\footnotesize\textcolor{AccentGray}{Stage: \theprfaqstage{} \enspace\textbar\enspace v\theprfaqmajor.\theprfaqminor}}
\fancyhead[R]{\footnotesize\textcolor{AccentGray}{\thepage}}
\renewcommand{\headrulewidth}{0pt}

% ══════════════════════════════════════════════════════════════════════════════
% Custom environments
% ══════════════════════════════════════════════════════════════════════════════

\newcommand{\prsection}[1]{%
  \vspace{1em}%
  {\large\bfseries\color{SectionBlue}#1}%
  \vspace{0.3em}\par%
}

\newmdenv[
  topline=false, bottomline=false, rightline=false,
  linewidth=3pt, linecolor=QuoteBorder, backgroundcolor=BoxBg,
  innerleftmargin=12pt, innerrightmargin=12pt,
  innertopmargin=8pt, innerbottommargin=8pt,
  skipabove=\baselineskip, skipbelow=\baselineskip,
]{customerquote}

\newmdenv[
  topline=false, bottomline=false, rightline=false,
  linewidth=3pt, linecolor=SectionBlue, backgroundcolor=BoxBg,
  innerleftmargin=12pt, innerrightmargin=12pt,
  innertopmargin=8pt, innerbottommargin=8pt,
  skipabove=\baselineskip, skipbelow=\baselineskip,
]{spokespersonquote}

\newcounter{faqnum}
\newenvironment{faqpair}[1]{%
  \needspace{4\baselineskip}%
  \vspace{0.5em}%
  \refstepcounter{faqnum}%
  \noindent\textbf{\color{SectionBlue}Q\thefaqnum: #1}\par\nopagebreak[4]%
  \vspace{0.2em}%
  \begin{adjustwidth}{1em}{0pt}%
  \setlength{\parindent}{0pt}%
}{%
  \end{adjustwidth}%
  \vspace{0.5em}%
}
\newcommand{\faqref}[1]{\hyperref[#1]{FAQ~\ref*{#1}}}

\newcounter{featurenum}
\newcommand{\featureitem}[2]{%
  \refstepcounter{featurenum}%
  \item[\textbf{\color{SectionBlue}F\thefeaturenum.}]%
  \textbf{#1} --- #2%
}
\newcommand{\featureref}[1]{\hyperref[#1]{Feature~\ref*{#1}}}

\usepackage{changepage}

% ══════════════════════════════════════════════════════════════════════════════
% Document
% ══════════════════════════════════════════════════════════════════════════════

\begin{document}

% ── Headline block ───────────────────────────────────────────────────────────
\begin{center}
  {\LARGE\bfseries\color{SectionBlue} Open-Source Tool Gives Human-Agent Teams\\Persistent Identity and Structured Delegation} \\[0.8em]
  {\large\color{AccentGray} Punt Labs releases ethos, a runtime that binds identity, typed mission contracts,} \\[0.3em]
  {\large\color{AccentGray} and role-based boundaries into one layer for humans and AI agents---} \\[0.3em]
  {\large\color{AccentGray} accessible via filesystem, CLI, or MCP, with zero vendor lock-in}
\end{center}

\vspace{0.5em}

% ══════════════════════════════════════════════════════════════════════════════
% PRESS RELEASE
% ══════════════════════════════════════════════════════════════════════════════

\section*{Press Release}

SAN FRANCISCO --- October 1, 2026 --- Punt Labs today announced ethos, a free, open-source runtime that gives human-agent engineering teams three capabilities no existing tool provides together: persistent identity for every participant, typed mission contracts for structured delegation, and a loose integration pattern that lets any tool---first-party or third-party---read identity and workflow state without taking a dependency (see \faqref{faq:differentiation}). Unlike agent persona conventions such as SoulSpec~\cite{soulspec2026standard}, which address agent-only identity, or orchestration frameworks such as CrewAI and Mastra, which coordinate agents without persistent identity or file-level ownership, ethos treats humans and agents as first-class team members with the same schema, the same delegation contracts, and the same audit trail. Whether developers want identity, delegation, and integration from one tool---rather than from separate specialized layers---is the central hypothesis this product tests (see \faqref{faq:customer-evidence}).

\prsection{Problem}

Today, when a developer opens an agentic coding tool, the session starts anonymous in both directions. The agent has no personality, no communication style, no team awareness. Sub-agents are interchangeable blanks. Every session begins from scratch~\cite{arxiv2025agenticmanifests,arxiv2025agentreadmes}.

This identity gap is only the first problem. The second is delegation chaos. When a leader assigns work to an agent, the contract is free-form prose: no file ownership, no round budget, no structured result, no audit trail. Two agents can edit the same files simultaneously. A reviewer can silently drift from the criteria it was pinned to at launch. When something goes wrong, the leader reconstructs events from chat history. An audit of 30 popular agent projects found 93\% use unscoped credentials with no scope narrowing, depth limits, or delegation chain controls~\cite{grantexagentsec2026}. The industry has agents that can write code but no discipline for assigning, bounding, and verifying the work they do (see \faqref{faq:customer-evidence}).

The third problem is tool fragmentation. A messaging tool reinvents usernames. A voice tool reinvents voice config. An email tool reinvents contact addresses. Each one solves a fragment of identity independently because no shared layer exists. SoulSpec~\cite{soulspec2026standard} addresses agent personas with structured markdown files, but only for agents, only for personality, and only where the host tool natively loads the files. A2A~\cite{a2av1milestone2026} provides cryptographic Agent Cards for discovery, but has no human identity model and no session persistence.

\prsection{Solution}

Ethos solves all three problems with one runtime. \textbf{Identity}: a developer runs {\small\texttt{ethos create}} once to define their persona---name, email, voice, writing style, personality---and creates agent personas the same way. Hooks load identity automatically at session start, survive context compaction, and inject matched personas when sub-agents spawn. Roles define tool restrictions and responsibilities; the team graph derives anti-responsibilities (``don't make architectural decisions---the COO handles that'') from {\small\texttt{reports\_to}} edges (\featureref{feat:teams}).

\textbf{Workflow}: the leader writes a typed mission contract---worker, write-set, success criteria, evaluator, round budget---and delegates. The store refuses overlapping write-sets, pins the evaluator's content hash at launch, and blocks close without a structured result for the current round. The same contract works for tight implementation (``one ReadFile, make check passes'') and open design (``evaluate three approaches, recommend one with tradeoffs''). The leader controls the agency level through contract specificity, not a mode switch (\featureref{feat:missions}).

\textbf{Integration}: any tool reads identity through the pattern that fits---YAML files at a known path (zero dependency), CLI commands from hooks and scripts, or the MCP server for structured operations. Three Punt Labs tools (Biff, Vox, Beadle) integrate this way today; any third-party tool can do the same. Extensions let tools attach per-persona config (voice profile, GPG key, routing preferences) without schema changes (\featureref{feat:extensions}).

\prsection{Customer Quote}

\begin{customerquote}
\textit{``I run a team of three engineers plus five agent personas---two Go specialists, a security reviewer, an architect, and a CLI designer. Before ethos, delegating to a sub-agent meant writing a paragraph of prose and hoping it stayed on task. Now I write a mission contract: which files to touch, what success looks like, and a cap of two rounds. The agent delivers a structured report that tells me what passed, what didn't, and where to look. The reviewer can't drift between rounds---if someone changes its criteria, the system catches it. A real bug fix went from task assignment to merged PR in under 13 minutes. I didn't make the delegation faster; the contract made it reliable.''}

\raggedleft --- \textbf{Priya Chakraborty}, Staff Engineer at a Series B startup
\end{customerquote}

\prsection{Getting Started}

Getting started takes under five minutes. First, install the binary with a single command:

{\small\texttt{curl -fsSL https://punt-labs.github.io/ethos/install.sh | sh}}

\noindent The installer seeds 6 starter roles, 10 domain talents, and the baseline-ops skill. Second, create your persona with
{\small\texttt{ethos create}}. If a
{\small\texttt{CLAUDE.md}} exists in the current project, the command detects it and fine-tunes starter content defaults---pre-filling writing style, personality, and role suggestions from the project context. 90\% of the result is batteries-included starter content; the remaining 10\% is LLM-inferred personalization from what the developer already has. The developer confirms or adjusts the suggestions and moves on. Create agent personas the same way. Third, add
{\small\texttt{agent: claude}} to your repo's
{\small\texttt{.punt-labs/ethos.yaml}}, optionally define a team with roles, and start a session. Hooks handle everything automatically. For structured delegation, write a mission contract and let the lifecycle drive the work.

\prsection{Spokesperson Quote}

\begin{spokespersonquote}
``The hardest design choice was making write-sets a convention, not an enforcement boundary. Agent runtimes don't support filesystem sandboxing today---if we enforced write-sets at the kernel level, we'd be building a sandbox product, not an identity product. So write-sets are declared in the contract, surfaced to the agent, and verified in review. That sounds weak until you see the alternative: every other tool gives agents unscoped access and hopes for the best. We also chose the same schema for humans and agents. That's not idealism---it's plumbing. When the voice tool reads a persona, it doesn't check whether the persona is human or machine. When a mission contract names an evaluator, the system pins the evaluator's content hash regardless of kind. One schema means every integration works for both, and we never maintain parallel paths.''

\raggedleft --- \textbf{Jim Freeman}, Founder at Punt Labs
\end{spokespersonquote}

\prsection{Call to Action}

Ethos is available now at
\url{https://github.com/punt-labs/ethos}
under the MIT license. It ships as a single Go binary with zero runtime dependencies. The Claude Code plugin is in the Punt Labs plugin marketplace. A live example walkthrough---real mission, real outputs, real timing---is at
{\small\texttt{docs/example/}} in the repository.

\newpage

% ══════════════════════════════════════════════════════════════════════════════
% FREQUENTLY ASKED QUESTIONS
% ══════════════════════════════════════════════════════════════════════════════

\section*{Frequently Asked Questions}

\subsection*{External FAQs}

\begin{faqpair}{What is ethos and who is it for?}\label{faq:what}
Ethos is an open-source runtime for engineering teams that work with AI agents. It provides three things: persistent identity (personality, writing style, talents, roles, team structure) that loads automatically every session; typed mission contracts for structured delegation with file-level ownership, bounded rounds, and audit trails; and a loose integration pattern where any tool reads identity state via filesystem, CLI, or MCP without taking a dependency. It is for developers who delegate real work to AI agents and need that delegation to be reliable, auditable, and bounded.
\end{faqpair}

\begin{faqpair}{How is ethos different from SoulSpec, CLAUDE.md, or agent orchestration frameworks?}\label{faq:differentiation}
SoulSpec~\cite{soulspec2026standard} provides structured agent personas (SOUL.md, IDENTITY.md, STYLE.md) adopted by OpenClaw (350,000+ GitHub stars)~\cite{arxiv2510soulmd}. It addresses agent-only personality with no human identity, no team model, no delegation contracts, and no cross-tool integration surface beyond file discovery. CLAUDE.md describes project conventions---coding standards, architecture decisions---not people~\cite{claudemd2025pnote}. Orchestration frameworks like CrewAI and Mastra~\cite{mastra2026supervisor} coordinate agents with roles and typed schemas but have no persistent identity across sessions, no file-level write-set boundaries, and no frozen evaluator to prevent criteria drift.

Ethos is the only tool that combines human-agent parity (same schema for both), typed delegation contracts (write-set, evaluator pinning, bounded rounds), and multi-surface integration (filesystem + CLI + MCP). The bet is that these three capabilities need to be one layer, not three separate tools.
\end{faqpair}

\begin{faqpair}{How do I get started?}
Install
({\small\texttt{curl -fsSL .../install.sh | sh}}),
run {\small\texttt{ethos create}} to define your persona, add
{\small\texttt{agent: claude}} to your repo's
{\small\texttt{.punt-labs/ethos.yaml}}, and start a session. Under five minutes. No account, no cloud, no API key.
\end{faqpair}

\begin{faqpair}{How much does it cost?}
Free and open-source under the MIT license. Ethos is infrastructure---designed for frictionless adoption, not direct monetization.
\end{faqpair}

\begin{faqpair}{What happens to my data?}\label{faq:data}
All data is local:
{\small\texttt{\textasciitilde/.punt-labs/ethos/}}.
No server, no telemetry, no cloud sync. Identity files are plain YAML, human-readable, editable with any text editor. Mission contracts and audit logs are local JSONL. The privacy surface is equivalent to
{\small\texttt{\textasciitilde/.gitconfig}}. 90\% of organizations see local storage as inherently safer, and 64\% worry about sharing sensitive data with cloud GenAI tools~\cite{cisco2025privacy}.
\end{faqpair}

\begin{faqpair}{What tools integrate with ethos today?}\label{faq:integrations}
Three integration patterns, any coupling level:

\begin{itemize}[nosep]
  \item \textbf{Filesystem.} Read YAML at known paths. Zero dependency.
  \item \textbf{CLI.} Call {\small\texttt{ethos whoami}}, {\small\texttt{ethos mission show}}, {\small\texttt{ethos session}} from hooks and scripts.
  \item \textbf{MCP server.} 10 tools covering identity, sessions, teams, roles, missions, and extensions.
\end{itemize}

Three Punt Labs tools (Biff for messaging, Vox for voice, Beadle for email) integrate today. Each works without ethos but gains richer identity context when it is present. The integration surface is open to any third-party tool---Zed's Agent Client Protocol~\cite{zedacp2026} is one example of a new interoperability surface ethos could target.
\end{faqpair}

\begin{faqpair}{Can I use missions for design work, not just implementation?}
Yes. The mission contract is a spectrum. An implementation mission has a tight write-set (``only this file'') and specific criteria (``one ReadFile, make check passes''). A design mission has a broad write-set (``docs/design/'') and open-ended criteria (``evaluate three approaches, recommend one with tradeoffs''). The agent's expertise and judgment are always present; the contract determines how much latitude they have. The leader chooses the structure level when writing the contract, not when defining the persona.
\end{faqpair}

\begin{faqpair}{Can I use ethos with SoulSpec or existing CLAUDE.md files?}\label{faq:interop}
Yes. Ethos both ingests and exports these formats.

\textbf{Ingestion.} Ethos reads SoulSpec files (SOUL.md, IDENTITY.md, STYLE.md, AGENTS.md) and CLAUDE.md as input during identity creation. The content becomes an ethos identity with structural metadata the source format cannot express---typed roles with tool restrictions, team collaboration graphs, mission contracts with write-set boundaries, safety constraints, and hash-pinned evaluators. The ingested identity is a hardened version of the same content: the personality is the same words, but now it is pinned by content hash so it cannot drift between mission rounds. The writing style is the same prose, but now it is bound to a role that restricts which tools the agent can use.

\textbf{Export.} Ethos exports to SoulSpec format (SOUL.md, IDENTITY.md, STYLE.md, AGENTS.md) for any compatible tool, and to CLAUDE.md for any Claude Code session without ethos installed. Both exports are lossy---structural enforcement (roles, teams, missions, audit trails, frozen evaluators) drops because markdown cannot represent it. The loss is the point: it shows what ethos adds. Export is useful for distribution (SoulSpec files work in OpenClaw) and portability (CLAUDE.md works without ethos).

\textbf{The relationship.} SoulSpec and CLAUDE.md are content formats---personality as prose, conventions as text. Ethos is the structural layer that makes that content enforceable. Import gains structure. Export loses structure but gains reach.
\end{faqpair}

\newpage
\subsection*{Internal FAQs}

\subsubsection*{Value \& Market}

\begin{faqpair}{What is the total addressable market?}\label{faq:tam}
\textbf{Bottoms-up estimate.} The immediate market is developers using terminal-based agentic coding tools: Claude Code, Codex CLI, OpenCode. Claude Code is the most-loved AI coding tool at 46\% (vs.\ Cursor 19\%, GitHub Copilot 9\%) per the Pragmatic Engineer survey ($n{\approx}906$)~\cite{pragmaticengsurvey2026}. 71\% of developers who regularly use AI agents use Claude Code. Anthropic's Claude Code run-rate is \$2.5B~\cite{demandsageclaude2026}. We estimate 200,000--500,000 developers in the terminal-first agentic segment based on Claude Code's 115,000 floor (July 2025)~\cite{claudecode2025stats} plus 2026 growth trajectory.

\textbf{Assumption:} Terminal-first segment size is inferred, not measured. The estimate assumes it grows faster than the broader market because multi-agent workflows are exclusively terminal-based today. Ethos is free; TAM is measured in adoption (installations, missions created), not revenue.
\end{faqpair}

\begin{faqpair}{What evidence do we have that customers want this?}\label{faq:customer-evidence}
Evidence is structural and observational. No customer interviews have been conducted.

\textbf{Identity gap:}
\begin{itemize}[nosep]
  \item An analysis of 253 CLAUDE.md files found they are ``primarily optimized for efficient code execution''---not for encoding who the user is~\cite{arxiv2025agenticmanifests}. A study of 2,303 context files across 1,925 repositories confirmed agents operate without identity context~\cite{arxiv2025agentreadmes}.
  \item SoulSpec emerged as a community standard for agent personas, adopted in 466 open-source projects~\cite{arxiv2510soulmd}. This confirms demand for structured identity---but only for agents, not humans, and without delegation or cross-tool integration.
  \item Claude Managed Agents (April 2026)~\cite{claudemanagedagents2026} uses ``persona'' for deployment configuration, not rich identity. Human developer identity is absent from the model.
\end{itemize}

\textbf{Delegation gap:}
\begin{itemize}[nosep]
  \item 93\% of popular agent projects use unscoped API keys with no delegation controls~\cite{grantexagentsec2026}. 48.9\% of organizations are entirely blind to machine-to-machine traffic~\cite{saltsecurity2026agentsec}.
  \item OWASP identifies Identity and Privilege Abuse as a top agentic AI risk: ``agents need their own identities with task-scoped, time-bound permissions and clear auditability''~\cite{owaspagentic2025}.
  \item Academic research confirms 2--3 reflection turns achieve unanimous convergence in multi-agent decision-making~\cite{arxiv2504consensus2026}---validating ethos's bounded-rounds-with-reflection design.
  \item NIST COSAiS (draft Q3 2026) targets least-privilege tool access, agent action containment, and chain-of-custody logging~\cite{nistcosais2026}---exactly the controls ethos's workflow pillar implements.
\end{itemize}

\textbf{Internal observation:} Three independently developed tools (messaging, voice, email) each reinvented a fragment of identity before a shared layer existed. One live mission (ethos-db7, April 2026) went from bead claim to merged PR in 12 minutes 55 seconds---17\% coding, 62\% review pipeline---demonstrating that the mission primitive works in practice.

\textbf{Validation plan:} Interview 10 developers who use Claude Code with sub-agents or Agent Teams. The discriminating question: is the combined identity + delegation gap painful enough to install a tool that fills both?
\end{faqpair}

\begin{faqpair}{Who are the competitors and why will we win?}\label{faq:competitors}
\begin{itemize}[nosep]
  \item \textbf{SoulSpec / OpenClaw}~\cite{soulspec2026standard,arxiv2510soulmd} --- structured agent personas (SOUL.md, IDENTITY.md, STYLE.md). Agent-only; no human identity, no teams, no delegation, no MCP/CLI surface. 350,000+ GitHub stars via OpenClaw. The closest identity competitor. Ethos ingests SoulSpec files as an onboarding path---SoulSpec users import their existing personas and gain structural enforcement (typed roles, team graphs, mission contracts) without starting from scratch. Ethos differentiates on human-agent parity, team graphs, and the workflow pillar.
  \item \textbf{Mastra}~\cite{mastra2026supervisor} --- typed Zod request schemas and pre-delegation hooks. 300,000+ weekly npm downloads. Closest to typed delegation. No write-set boundaries, no frozen evaluator, no append-only audit, no identity layer.
  \item \textbf{Claude Managed Agents}~\cite{claudemanagedagents2026} --- Anthropic's hosted agent infrastructure (April 2026). Stateful sessions, scoped permissions. ``Persona'' means deployment configuration, not rich identity. No human developer identity. Vendor-specific.
  \item \textbf{A2A v1.0}~\cite{a2av1milestone2026} --- 150+ organizations, Signed Agent Cards for cryptographic identity. Agent discovery protocol, not a persona or delegation layer. No human identity model.
  \item \textbf{CLAUDE.md / AGENTS.md}~\cite{claudemd2025pnote} --- per-project conventions, not per-person identity. Complementary.
  \item \textbf{CrewAI} --- role-based agent orchestration. Natural-language delegation, no typed contracts, no persistent identity, no write-set or reflection.
  \item \textbf{Entire.io}~\cite{entire2026techcrunch,entireio2026docs,entireio2026hello} --- Thomas Dohmke (ex-GitHub CEO), \$60M seed at \$300M valuation. Entire captures session fidelity---every token, tool call, and prompt---as retrospective checkpoints stored on a dedicated git branch ({\small\texttt{entire/checkpoints/v1}}). AI-generated summaries (intent, outcome, learnings, friction, open items) attach to commits after the session ends. The critical difference: Entire records what the AI did (retrospective forensics); ethos specifies what the AI should do before it acts (prospective contracts). Entire has no pre-delegation spec, no write-set boundaries, no typed mission lifecycle, no ADRs. Ethos has no full-transcript capture---that role is filled by Quarry (optional). The two are complementary: Entire answers ``what did the AI do?'' Ethos answers ``what did we ask it to do, why, and did it stay in bounds?''
  \item \textbf{gstack}~\cite{gstack2025github} --- 23 SKILL.md prompt files by Garry Tan (YC CEO) simulating an engineering sprint team. Not a competitor but a validation: gstack proves developers want sprint discipline for agent teams. The content is high-quality workflow design with no structural enforcement---write-sets are suggestions, safety is conversational, personas reset each invocation. Ethos provides the enforcement layer that makes prompt-based workflows structural. Complementary: a gstack-inspired sprint team ships as a batteries-included ethos starter template.
\end{itemize}

\textbf{Why ethos wins:} No competitor combines human-agent identity parity, typed delegation contracts with file-level ownership, and a multi-surface integration pattern. SoulSpec has identity (agent-only). Mastra has delegation hooks (no identity). A2A has discovery (no personas, no delegation). Ethos is the only tool where the same runtime gives an agent its personality, constrains its file access in a mission, and lets a voice tool read its voice config---all from the same identity record.

\textbf{Competitive response:} If Anthropic ships native human identity in Claude Code, the identity pillar becomes less necessary as a standalone layer. The workflow pillar (missions, write-sets, audit trails) and the cross-tool integration pattern would retain independent value. Entire.io validates the governance market at a \$300M valuation~\cite{entire2026techcrunch}, confirming that intent capture and audit trails for agent-written code are a real category. Ethos's approach is prospective (contracts before delegation) rather than retrospective (checkpoints after commits), and built-in rather than a separate product. The bet is that the agentic tool market remains fragmented and that identity + delegation + integration need to be one layer.
\end{faqpair}

\begin{faqpair}{What is the customer acquisition strategy?}\label{faq:acquisition}
Five channels, ordered by expected contribution:

\begin{enumerate}[nosep]
  \item \textbf{Organic search and GitHub discovery.} Developers searching for ``Claude Code identity,'' ``agent delegation,'' or ``AI team structure'' find the repo. The MIT license and zero-dependency install remove friction. README, SEO-targeted docs, and the live example walkthrough convert visitors to installers.
  \item \textbf{Punt Labs tool ecosystem.} Biff, Vox, and Beadle surface ethos during normal use---``install ethos for richer identity context'' prompts appear when the tool detects ethos is absent. Each tool is a distribution channel for the shared layer.
  \item \textbf{Community and content.} Blog posts, conference talks, and example repos demonstrating mission contracts and team structures. The 12m55s end-to-end demo is the core asset: a real bug fix from bead claim to merged PR, timed and reproducible.
  \item \textbf{Portable export and ecosystem interoperability.} Lossy export (\featureref{feat:lossy-export}) lets ethos users publish identities into two ecosystems: SoulSpec files (SOUL.md, IDENTITY.md, STYLE.md, AGENTS.md) for OpenClaw and compatible tools, and CLAUDE.md personal sections for any Claude Code session without ethos installed. A2A-compatible discovery lets ethos identities participate in cross-platform agent networks. Developers in those ecosystems discover ethos when they want human identity parity, team graphs, or delegation contracts that their current tools lack.
  \item \textbf{Content compatibility.} Ethos imports structured workflows from popular prompt-based tools. gstack's 23 SKILL.md files~\cite{gstack2025github} become 6 identities with enforced roles; SoulSpec personas map to ethos identities with team structure. Users of these tools discover ethos when they want structural enforcement behind their prompt-based workflows: ``you already have the workflow discipline as prompts---ethos makes it structural.'' A gstack-inspired sprint team ships as a batteries-included starter template, so developers start with a working team structure on day one.
\end{enumerate}

\textbf{Assumption:} Channel contribution is judgment-based. No acquisition data exists. The validation interviews will test whether ``found it on GitHub'' or ``heard about it from a tool prompt'' is the dominant path.
\end{faqpair}

\begin{faqpair}{What is your next step to validate your vision?}
Interview 10 developers who use Claude Code with sub-agents or Agent Teams. The discriminating question: \textbf{is the combined identity + delegation gap painful enough to install a tool that fills both?}

The interview protocol:
\begin{enumerate}[nosep]
  \item Do you configure personal preferences in CLAUDE.md? How many repos?
  \item When you delegate to a sub-agent, how do you specify what files it can touch? What happens when two agents edit the same file?
  \item \textbf{Demo:} Show (a) session where the agent greets the developer by name with a distinct personality, (b) a mission contract that bounds a delegation to specific files with 2 rounds max, (c) the audit trail after the mission closes.
  \item After seeing this: would you install a tool that provides identity + delegation structure, knowing it takes 5 minutes and stores everything locally?
  \item If Claude Code added native identity and delegation next month, would you still want a cross-tool layer?
\end{enumerate}

Success threshold: 7 of 10 describe the combined gap as a daily friction or would install ethos after the demo. If most say ``nice but I'd wait for Claude Code,'' the gap is real but the urgency is insufficient for a standalone tool.
\end{faqpair}

\subsubsection*{Technical}

\begin{faqpair}{What are the major technical risks?}\label{faq:architecture}
\textbf{Risk 1: Integration surface adoption.} Three integration patterns (filesystem, CLI, MCP) require documentation and examples. Third-party adoption is unproven. Mitigation: three first-party tools (Biff, Vox, Beadle) demonstrate the patterns. Documentation ships with the binary.

\textbf{Risk 2: Mission contract complexity.} The workflow pillar has 7 primitives (contract, write-set, frozen evaluator, bounded rounds, verifier isolation, result artifacts, event log). New users may find the full lifecycle overwhelming. Mitigation: the mission is optional---identity works without it. The spectrum from tight implementation to open design means users can start simple. The live example (12m55s, docs/example/) demonstrates the full flow.

\textbf{Risk 3: Schema evolution.} Core identity fields are fixed; extensions live in separate files per tool. Mission contracts use strict unknown-field-rejecting decode. Adding fields requires careful versioning. Mitigation: the extension model handles per-tool data without core schema changes. The strict decoder prevents silent corruption.

\textbf{Risk 4: SoulSpec competition.} SoulSpec is gaining traction (466 OSS projects, OpenClaw adoption). If SoulSpec expands to include human identity and cross-tool integration, ethos's identity pillar faces direct competition. Mitigation: the workflow pillar and multi-surface integration are independent of the identity competition. Lossy export to SoulSpec and CLAUDE.md formats (\featureref{feat:lossy-export}) turns this into coexistence rather than competition---ethos ingests SoulSpec files on the way in and exports them on the way out.
\end{faqpair}

\begin{faqpair}{What dependencies exist on other teams or systems?}
None. Ethos is a standalone Go binary with zero runtime dependencies. It does not depend on Biff, Vox, Beadle, or any external service. The reverse dependency is graduated: filesystem readers have no binary dependency; CLI and MCP callers need the binary installed. If ethos is absent, tools fall back to their own identity resolution.
\end{faqpair}

\begin{faqpair}{What is the development status?}
Ethos has shipped 4 of 5 planned phases. 20,100 lines of production Go. 26,800 lines of tests (1.33:1 ratio). Go Report Card: A+.

\begin{itemize}[nosep]
  \item \textbf{Phase 1 (Shipped):} Identity, sessions, persona animation, teams, roles, 10 talents, 6 role archetypes, baseline-ops skill.
  \item \textbf{Phase 2 (Shipped):} Anti-responsibility generation, role-based hooks, structured output, baseline-ops injection, mission skill MVP.
  \item \textbf{Phase 3 (Shipped):} All 7 workflow primitives: mission contract, write-set admission, frozen evaluator, bounded rounds with reflection, verifier isolation, result artifacts, append-only event log. 7 ADRs (DES-031 through DES-037).
  \item \textbf{Phase 4 (Shipped):} SessionStart working context (git branch, dirty files), role-based safety constraints, session audit logging.
  \item \textbf{Phase 5 (Planned):} Starter team templates, cross-tool integration (ethos+biff, ethos+beadle), agent marketplace.
\end{itemize}

Remaining work is adoption-driven: customer validation, integration guides for third-party tools, and Phase 5 ecosystem features.
\end{faqpair}

\begin{faqpair}{What is the scaling story?}
Ethos is a local tool---no server, no database. Scaling means identities on one machine and concurrent sessions reading them. At 100 identities and 50 concurrent sessions, the bottleneck is filesystem I/O for flock contention, which Go handles efficiently. Mission state is session-scoped. No user will realistically hit these limits.
\end{faqpair}

\begin{faqpair}{How does ethos capture intent and maintain audit trails?}\label{faq:audit-trails}
Three layers, each independent:

\textbf{Git (durable, shared).} Mission contracts, structured results, and typed ADRs live in {\small\texttt{.ethos/}} in the repository. The contract captures intent before delegation---what was specified, what files are owned, what success looks like. The result captures outcome---what was delivered, what passed, what didn't. ADRs capture design decisions with a status lifecycle (proposed $\to$ settled $\to$ superseded), rejected alternatives, and links to the missions where decisions were made~\cite{promptpals2026adrsai}. All travel with the code in version control.

\textbf{Local (operational).} Event-by-event JSONL logs per mission and per-session tool audit trails stay on the developer's machine. Useful for debugging and post-mortems but not shared. This is operational telemetry, not the permanent record.

\textbf{Quarry (optional, semantic).} When Quarry is installed, conversation transcripts are captured before context compaction and tagged with the active mission ID. Semantic search across sessions finds the reasoning behind decisions---``why did we choose conventions over enforcement?'' retrieves the conversation where DES-041 was discussed. Quarry adds the unstructured ``why'' layer that structured artifacts cannot capture.

Together: git answers ``what was decided,'' local logs answer ``what happened in detail,'' Quarry answers ``why was it decided.'' Each layer works independently. This is a lightweight built-in alternative to Entire.io's checkpoint-branch approach~\cite{entireio2026docs}---zero additional infrastructure for teams already using git. Entire captures the full transcript retrospectively; ethos captures intent prospectively in the contract and outcome structurally in the result. The two approaches are complementary rather than competing.
\end{faqpair}

\subsubsection*{Business}

\begin{faqpair}{What is the revenue model?}
Free and open-source. The strategic value is adoption-driven: wider ethos adoption increases the value of every tool built on the identity and workflow layer. When a developer installs ethos once, messaging, voice, email, code review, and any future tool gain richer context without reimplementing identity or delegation.
\end{faqpair}

\begin{faqpair}{What does the P\&L look like at steady state?}\label{faq:pnl}
Ethos has no revenue line. The P\&L is a cost analysis.

\textbf{Costs (annual):} Engineering maintenance: 40--80 hours/year (bug fixes, Go upgrades, schema evolution). CI/CD: negligible (\textless\$50/year). No infrastructure cost.

\textbf{Strategic value (assumed):} If any downstream tool reaches 200 users and ethos contributes a 10\% retention improvement, the value exceeds maintenance cost. Testable once downstream tools have users.

\textbf{Decision threshold:} If after 12 months ethos has fewer than 100 installations and no measurable lift to downstream tool retention, fold identity and workflow into each consumer tool and deprecate the standalone binary.
\end{faqpair}

\begin{faqpair}{What are the key metrics?}
\begin{enumerate}[nosep]
  \item \textbf{Installations}: 500 in 6 months (GitHub release downloads).
  \item \textbf{Missions created}: evidence that the workflow pillar is used, not just identity. Target: 100 missions in 6 months.
  \item \textbf{Tool integrations}: 4 within 6 months (first-party + third-party).
  \item \textbf{Third-party extensions}: 1 within 12 months.
\end{enumerate}

Decision threshold: installations below 100 after 6 months = identity layer is not a pull product. Missions below 10 = workflow pillar is not valued. Either triggers reconsideration.

\textbf{Reference class note:} The 500-install and 100-mission targets are judgment-based. No direct reference class exists for a free, open-source identity-and-workflow tool in this product category. The decision thresholds (100 installs, 10 missions) are the actionable numbers---they determine whether ethos continues as a standalone product. The higher targets are aspirational indicators, not kill criteria.
\end{faqpair}

\begin{faqpair}{What are we not building?}\label{faq:not-building}
Two strategic exclusions deserve explanation (see also \featureref{feat:no-runtime-enforce} and \featureref{feat:no-remote} in the Feature Appendix):

\textbf{Runtime enforcement.} Write-sets are conventions: declared in the mission contract, surfaced to the agent, and verified in review. Ethos does not enforce them at the filesystem level. This is deliberate. Agent runtimes (Claude Code, Codex CLI) do not expose a sandboxing API today. Building kernel-level containment would mean shipping a sandbox product, not an identity product---a different problem with different customers. If agent runtimes add scoped filesystem access in the future, ethos can emit the constraints for them to enforce. Until then, convention-plus-review is the only mechanism that works across all runtimes.

\textbf{Cloud sync.} All identity and mission state is local. This blocks the most obvious enterprise adoption path (centralized team management, SSO-provisioned identities) but keeps the tool simple, private, and zero-dependency. The tradeoff is intentional: local-only means no authentication, no trust model, no server infrastructure, and no data privacy concerns. If enterprise demand materializes and requires centralized identity, the right answer is a separate sync layer that reads ethos's local state---not cloud features grafted onto the core binary.
\end{faqpair}

\begin{faqpair}{Why now?}
Five shifts converged:

\begin{enumerate}[nosep]
  \item \textbf{Agent coding went mainstream.} Claude Code is the most-loved AI coding tool (46\%, $n{\approx}906$)~\cite{pragmaticengsurvey2026}. 84\% of developers use or plan to use AI tools~\cite{stackoverflow2025survey}. Multi-agent sessions (Agent Teams, sub-agents) created 3--10 agent scenarios where anonymous coordination breaks.
  \item \textbf{The delegation gap became a security problem.} 93\% of agent projects use unscoped credentials~\cite{grantexagentsec2026}. 48.9\% of organizations are blind to machine-to-machine traffic~\cite{saltsecurity2026agentsec}. NIST is drafting agent governance standards for 2027~\cite{nistcosais2026}. Ethos ships the controls ahead of the standard.
  \item \textbf{Agent identity became a real category.} SoulSpec launched as an open standard~\cite{soulspec2026standard}. Claude Managed Agents uses ``persona'' in its model~\cite{claudemanagedagents2026}. A2A v1.0 introduced Signed Agent Cards~\cite{a2av1milestone2026}. The vocabulary and demand exist---but no tool addresses humans, delegation, and cross-tool integration together.
  \item \textbf{The community built workarounds.} SOUL.md, IDENTITY.md, USER.md appeared as informal conventions~\cite{blueoctopus2025soulmd}. None standardized, none portable, none covering delegation. Ethos provides the structured, portable layer these workarounds approximate.
  \item \textbf{The market is growing fast enough to support new infrastructure.} The AI agent market is \$7.63B in 2025, growing at 45.8\% CAGR to \$50.31B by 2030~\cite{aidemandsageagents2026}. 79\% of enterprises have adopted AI agents, but only 11\% are in production---the adoption-production gap is where governance tools like ethos create value. Gartner projects 40\% of enterprise apps will embed agents by end of 2026, but warns over 40\% of agentic AI projects risk cancellation without governance~\cite{gartner2025multiagentsurge}.
\end{enumerate}
\end{faqpair}

% ══════════════════════════════════════════════════════════════════════════════
% FOUR RISKS ASSESSMENT
% ══════════════════════════════════════════════════════════════════════════════

\newpage
\section*{Risk Assessment}

\renewcommand{\arraystretch}{1.6}
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{\color{SectionBlue}Risk} & \textbf{\color{SectionBlue}Rating} & \textbf{\color{SectionBlue}Assessment} \\
\midrule
Value      & \textcolor{RiskAmber}{Medium} & The identity gap is structurally confirmed (253 CLAUDE.md files, 466 SoulSpec projects, OWASP top 10). The delegation gap is confirmed (93\% unscoped credentials, 48.9\% blind to M2M traffic). The combined product---identity + workflow + integration in one layer---is untested with customers. SoulSpec validates identity demand but only for agents. Missions validated in practice (12m55s end-to-end) but not by external users. The ``pull'' signal is untested: do developers want a unified layer, or do they expect identity and delegation from separate tools? \\
Usability  & \textcolor{RiskAmber}{Medium} & Identity onboarding is simple (three commands, five minutes). The mission workflow introduces 7 primitives that have not been tested with external users. The live example demonstrates feasibility, not usability---it was operated by the builder. \\
Feasibility & \textcolor{RiskGreen}{Low} & 4 of 5 phases shipped. 20.1 KLOC production Go, 26.8 KLOC tests, A+ Go Report Card. All 7 workflow primitives, 4 integration hooks, audit logging, and safety constraints are implemented and tested. No unsolved technical problems. \\
Viability  & \textcolor{RiskAmber}{Medium} & Ethos is free infrastructure whose viability depends on adoption and downstream tool value. Maintenance cost is low (40--80 hrs/year). If SoulSpec captures agent identity and a dominant tool ships native delegation, both pillars face competition. The cross-tool integration pattern is the most defensible value---it is the only structural advantage that requires a shared layer rather than per-tool features. \\
\bottomrule
\end{tabularx}
\renewcommand{\arraystretch}{1.0}

\prsection{Pre-Mortem}

Three scenarios would cause ethos to fail:

\begin{enumerate}[nosep]
  \item \textbf{SoulSpec wins agent identity; a dominant tool ships delegation.} If SoulSpec becomes the standard for agent personas and Claude Code ships native mission contracts, ethos is squeezed from both sides. The cross-tool integration pattern survives only if the market stays fragmented. Mitigation: ethos ingests SoulSpec files and exports to both SoulSpec and CLAUDE.md formats---interoperating rather than competing.
  \item \textbf{The combined product is too broad.} Identity + workflow + integration may be too much for one tool. Developers may want ``just identity'' or ``just delegation'' from focused tools. Mitigation: each pillar works independently. Identity loads without missions. Missions work without team structure. The pillars compound but don't require each other.
  \item \textbf{The benefit is too abstract to motivate installation.} Developers may understand the gaps intellectually but not feel them strongly enough to install a tool---especially when Claude Code adds native features quarterly. Validation interviews will test this. If 7 of 10 say ``I'd wait for native,'' the standalone tool should be reconsidered.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
% FEATURE APPENDIX
% ══════════════════════════════════════════════════════════════════════════════

\newpage
\section*{Feature Appendix}

\subsection*{Must Do (Shipped)}

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{Identity CRUD}{create, read, update, delete personas---same schema for humans and agents}\label{feat:crud}
  \featureitem{Persona animation}{SessionStart, PreCompact, and SubagentStart hooks inject personality, writing style, and talents automatically}\label{feat:animation}
  \featureitem{Teams and roles}{team graph with collaboration edges, role-based tool restrictions, anti-responsibility derivation from reports\_to edges}\label{feat:teams}
  \featureitem{Agent file generation}{SessionStart generates .claude/agents/<handle>.md from identity, personality, role, and team data}\label{feat:agentgen}
  \featureitem{Mission contracts}{typed delegation with leader, worker, evaluator, write-set, success criteria, and round budget}\label{feat:missions}
  \featureitem{Write-set admission}{segment-prefix conflict scan refuses overlapping file claims between open missions}\label{feat:writeset}
  \featureitem{Frozen evaluator}{sha256 content hash pinned at mission launch; SubagentStart refuses verifier spawns if the hash drifts}\label{feat:frozen}
  \featureitem{Bounded rounds with reflection}{hard round cap with mandatory structured reflection before advancing}\label{feat:rounds}
  \featureitem{Verifier isolation}{verifier subagent receives only the contract and file allowlist, not the worker's scratch state}\label{feat:isolation}
  \featureitem{Result artifacts and close gate}{typed results with verdict, confidence, files\_changed, evidence; close refused without a valid result}\label{feat:results}
  \featureitem{Append-only event log}{JSONL audit trail per mission; one corrupt line does not erase the tail}\label{feat:eventlog}
  \featureitem{SessionStart working context}{git branch, uncommitted changes, dirty file count injected at session start}\label{feat:workctx}
  \featureitem{Role-based safety constraints}{tool + message pairs emitted in generated agent files}\label{feat:safety}
  \featureitem{Session audit logging}{PostToolUse hook appends one JSONL line per tool invocation to a per-session audit log}\label{feat:auditlog}
  \featureitem{Three integration surfaces}{filesystem (zero dependency), CLI (hooks/scripts), MCP server (10 tools)}\label{feat:surfaces}
  \featureitem{Extension model}{per-tool attributes under <handle>.ext/<tool>.yaml; any tool attaches config without schema changes}\label{feat:extensions}
  \featureitem{Starter content}{10 domain talents, 6 role archetypes, baseline-ops skill ship in the box}\label{feat:starter}
\end{enumerate}

\subsection*{Should Do}

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{Mission archetypes}{named contract conventions---implement (tight write-set, specific criteria, agent executes), design (broad write-set, open criteria, agent proposes and pushes back), review (read-only, findings with confidence), investigate (read-only, questions answered). The agent's expertise is always present; the archetype determines how much latitude the contract gives. Agency is expressed through contract specificity, not a mode switch}\label{feat:archetypes}
  \featureitem{Lossy export to portable formats}{{\small\texttt{ethos export --to soulspec}} produces SOUL.md, IDENTITY.md, STYLE.md, AGENTS.md readable by any SoulSpec-compatible tool. {\small\texttt{ethos export --to claude-md}} produces a CLAUDE.md personal section usable without ethos installed. Both exports are lossy by design: structural data (teams, missions, safety constraints, frozen evaluator, audit trails) drops because the target format cannot represent it. The loss signals what ethos adds. Export is compilation from a richer system to a simpler one---useful for distribution and portability}\label{feat:lossy-export}
  \featureitem{Integration guides}{documentation and examples for third-party tools to read ethos identity and mission state}\label{feat:guides}
  \featureitem{Starter team templates}{solo developer (1 human + 3 agents), small team (2--3 humans + 5 agents), reference implementation}\label{feat:templates}
  \featureitem{LLM-assisted identity creation from project context}{{\small\texttt{ethos create}} reads CLAUDE.md and fine-tunes starter content defaults with LLM inference---pre-filling writing style, personality, and role suggestions from existing project conventions}\label{feat:llm-create}
  \featureitem{Structured workflow importers}{{\small\texttt{ethos import --from gstack|soulspec|crewai}} extracts personas and roles from popular tool formats; ships a gstack-inspired sprint team as starter content}\label{feat:importers}
  \featureitem{Git-tracked mission artifacts}{mission contracts and structured results committed to {\small\texttt{.ethos/}} in the repository. Intent and outcome travel with the code. Lightweight alternative to Entire.io's checkpoint-branch approach~\cite{entireio2026docs}---uses conventional git commits, not a separate branch structure}\label{feat:git-missions}
  \featureitem{Typed ADRs}{Architecture Decision Records as ethos artifacts with schema validation, lifecycle status (proposed/settled/superseded), and links to missions and beads~\cite{promptpals2026adrsai}. The agent reads ADRs before acting; the ADR captures the design decision that governed the delegation}\label{feat:typed-adrs}
  \featureitem{Quarry transcript tagging}{when Quarry is installed, conversation transcripts captured before context compaction are tagged with the active mission ID. Semantic search retrieves the reasoning behind specific missions---the unstructured ``why'' layer that completes the three-layer audit model}\label{feat:quarry-tag}
\end{enumerate}

\subsection*{Won't Do}

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{Cloud sync or remote identity}{ethos is local-only. Remote identity introduces trust, authentication, and privacy concerns that are out of scope for this layer.}\label{feat:no-remote}
  \featureitem{Authentication or authorization}{ethos provides identity (who is this?), not access control (what can they do?). A different layer.}\label{feat:no-auth}
  \featureitem{Consumer-specific fields}{the extension model exists for per-tool data. Ethos will never add fields for one consumer.}\label{feat:no-consumer}
  \featureitem{Runtime delegation enforcement}{write-set and safety constraints are surfaced to the agent and verified in review, not enforced at the filesystem level. Kernel-level sandboxing is a different product.}\label{feat:no-runtime-enforce}
  \featureitem{Telemetry or analytics}{no usage data collected. Metrics measured externally.}\label{feat:no-telemetry}
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
% REFERENCES
% ══════════════════════════════════════════════════════════════════════════════

\newpage
{\emergencystretch=5em\printbibliography[title={References}]}

\end{document}
