%
% architecture.tex
%
% Ethos: Technical Architecture Specification
% Written for engineers integrating identity, workflow, and
% cross-tool integration into AI agent systems.
%
% Compile: pdflatex architecture.tex
%

\documentclass[a4paper,10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\lstset{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue!80!black},
  commentstyle=\color{green!50!black},
  stringstyle=\color{red!60!black},
  breaklines=true,
  frame=single,
  framerule=0.3pt,
  rulecolor=\color{gray!50},
  backgroundcolor=\color{gray!5},
  columns=fullflexible,
}

\newcommand{\pkg}[1]{\texttt{#1}}
\newcommand{\tool}[1]{\texttt{#1}}
\newcommand{\field}[1]{\texttt{#1}}
\newcommand{\cmd}[1]{\texttt{#1}}
\newcommand{\fpath}[1]{\texttt{#1}}

\begin{document}

\title{Ethos: Technical Architecture Specification}
\author{Identity, Workflow, and Integration for Humans and AI Agents}
\date{April 2026 --- Phase 3 complete}
\hypersetup{
  pdftitle={Ethos: Technical Architecture Specification},
  pdfauthor={Punt Labs},
  pdfsubject={Architecture},
  pdfcreator={pdflatex}
}
\maketitle

\tableofcontents
\newpage

% ===================================================================
\section{Overview}
% ===================================================================

Ethos is the runtime that makes AI agents first-class citizens of an
engineering organization.  It binds three pillars into a single
system: a durable \emph{identity} (who the agent is), a typed
\emph{workflow} contract (how the agent works), and cross-tool
\emph{integration} (how the agent participates in email, chat, code
review, and voice).  Each pillar reinforces the others.  The identity
is the anchor; the mission contract is the operational primitive
that turns delegation into a runtime artifact; the integration layer
makes the same agent show up correctly in every channel.

\paragraph{Pillar 1 --- Identity.} One YAML file per persona.  Same
schema for humans and agents; the \field{kind} field is the only
structural difference.  An identity binds a persona (personality,
writing style, talents), a role (tools, responsibilities,
anti-responsibilities, team position), and channel bindings (email,
GitHub, voice, agent definition).  Identity data lives in two
layers: repo-local (git-tracked via the \fpath{.punt-labs/ethos/}
submodule) and user-global (\fpath{\~{}/.punt-labs/ethos/}).  Repo
takes priority for identity, attribute, role, and team lookup;
extensions always resolve from the global layer.

\paragraph{Pillar 2 --- Workflow.} The mission primitive turns
delegation from free-form prose into a typed runtime contract per
DES-031.  A mission contract carries a write-set (the file allowlist
the worker may touch), a frozen evaluator (pinned at launch,
content-hashed per DES-033), a round budget (bounded, with mandatory
reflection between rounds per DES-034), typed result artifacts
(schema-validated, append-only sibling files per DES-036), and an
append-only JSONL event log (every state transition, post-mortem
auditable per DES-037).  The store refuses operations that violate
the contract.  CLI and MCP surfaces are symmetric: every gate fires
identically regardless of how the contract reached the store.

\paragraph{Pillar 3 --- Integration.} Agent definition files in
\fpath{.claude/agents/<handle>.md} are generated from identity data,
augmented with anti-responsibilities derived from the team
collaboration graph, role-based behavioral hooks, structured output
formats, and the \texttt{baseline-ops} skill.  Operational hooks
inject working context at session start, enforce role-based
pre-tool safety constraints, and audit every tool invocation.
Cross-tool integration means an agent impersonating the COO sends
email from \texttt{claude@punt-labs.com} with the right GPG key
(Beadle), signs commits as \texttt{Claude Agento} with the right
SSH key (Biff), and speaks with the right voice (Vox) --- the
identity binds the channels and each tool reads ethos state to
configure itself.

Ethos is written in Go and ships as a CLI binary plus a Claude Code
plugin with an MCP server.  It is a \emph{sidecar}, not a
dependency.  Other tools (Vox, Beadle, Biff) read ethos state from
known filesystem paths.  They do not import ethos.  The file format
is the contract.

% ===================================================================
\section{Status at a Glance}
% ===================================================================

Ethos is built in phases.  Phase 1 (the identity layer), Phase 2.1--2.5
(production-quality agent generation), and Phase 3 (the workflow
primitives) are shipped; Phase 2.6 / Phase B--C (mission skill conflict
detection and dry-run) remains planned.  Phase 4 (operational hooks) and the
cross-tool integration deepening are planned.  This section is the
single authoritative status statement; the remainder of the document
describes the system in present tense as a coherent whole.

\begin{center}
\begin{longtable}{@{}lllll@{}}
\toprule
\textbf{Subsystem} & \textbf{Phase} & \textbf{ADR} & \textbf{Bead} &
  \textbf{Status} \\
\midrule
\endfirsthead
\toprule
\textbf{Subsystem} & \textbf{Phase} & \textbf{ADR} & \textbf{Bead} &
  \textbf{Status} \\
\midrule
\endhead
Identity registry              & 1   & DES-001--019 & ---            & SHIPPED \\
Attribute system               & 1   & DES-005      & ---            & SHIPPED \\
Extension system               & 1   & DES-009, 022 & ---            & SHIPPED \\
Session roster                 & 1   & DES-007, 023 & ---            & SHIPPED \\
Process tree walker            & 1   & DES-014      & ---            & SHIPPED \\
Layered store + resolution     & 1   & DES-013, 016 & ---            & SHIPPED \\
Persona animation hooks        & 1   & DES-008, 015 & ---            & SHIPPED \\
Teams and roles                & 1   & DES-024--028 & ---            & SHIPPED \\
Agent file generation          & 1   & DES-026      & ---            & SHIPPED \\
Starter content (10 talents,
  6 roles, baseline-ops skill) & 1   & DES-029, 030 & \texttt{ethos-l9d} & SHIPPED \\
\midrule
Anti-responsibility derivation & 2.1 & --- & \texttt{ethos-9ai.1} & SHIPPED \\
Role-based hook frontmatter    & 2.2 & --- & \texttt{ethos-9ai.2} & SHIPPED \\
Structured output for handoff  & 2.3 & --- & \texttt{ethos-9ai.3} & SHIPPED \\
Baseline-ops skill injection   & 2.4 & --- & \texttt{ethos-9ai.4} & SHIPPED \\
Mission skill Phase A (\cmd{/mission}) & 2.5 & --- & \texttt{ethos-9ai.5} & SHIPPED \\
Mission skill Phase B--C               & 2.6 & --- & \texttt{ethos-9ai.5} & PLANNED \\
\midrule
Mission contract               & 3.1 & DES-031      & \texttt{ethos-07m.5} & SHIPPED \\
Write-set admission control    & 3.2 & DES-032      & \texttt{ethos-07m.6} & SHIPPED \\
Frozen evaluator (content hash)& 3.3 & DES-033      & \texttt{ethos-07m.7} & SHIPPED \\
Bounded rounds + reflection    & 3.4 & DES-034      & \texttt{ethos-07m.8} & SHIPPED \\
Verifier isolation             & 3.5 & DES-035      & \texttt{ethos-07m.9} & SHIPPED \\
Result artifacts + close gate  & 3.6 & DES-036      & \texttt{ethos-07m.10} & SHIPPED \\
Event log reader API           & 3.7 & DES-037      & \texttt{ethos-07m.11} & SHIPPED \\
\midrule
RunE handler refactor          & Test & DES-042      & \texttt{ethos-aeb}   & SHIPPED \\
CLI subprocess tests (L2)      & Test & ---          & \texttt{ethos-oto}   & SHIPPED \\
In-process handler tests (L2.5)& Test & DES-042      & \texttt{ethos-a1m}   & SHIPPED \\
MCP integration tests (L3)     & Test & ---          & \texttt{ethos-764}   & SHIPPED \\
Behavioral test harness (L4)   & Test & DES-043      & \texttt{ethos-fal}   & SHIPPED \\
Ext layered resolution fix     & Test & DES-044      & \texttt{ethos-q6q}   & SHIPPED \\
\midrule
SessionStart working context   & 4.1 & ---          & \texttt{ethos-gcq.1} & PLANNED \\
Role-based pre-tool safety     & 4.2 & ---          & \texttt{ethos-gcq.2} & PLANNED \\
PostToolUse audit logging      & 4.3 & ---          & \texttt{ethos-gcq.3} & PLANNED \\
\midrule
Biff identity-aware messaging  & Integration & ---  & \texttt{ethos-wb4} & PLANNED \\
Beadle identity-aware email    & Integration & ---  & \texttt{ethos-g2f} & PLANNED \\
Vox voice config via extension & Integration & ---  & ---            & SHIPPED \\
\midrule
Starter team templates         & 5.1 & ---          & ---            & FUTURE \\
Agent marketplace              & 5.2 & ---          & ---            & FUTURE \\
Cross-tool orchestration       & 5.3 & ---          & ---            & FUTURE \\
\bottomrule
\end{longtable}
\end{center}

\textbf{Phase 3 complete.} After DES-037 shipped, the four
architecture rules from \texttt{\~{}/Documents/agents-architecture.tex}
are runtime-enforced for the first time in the project's history:
roles are interfaces, understanding centralizes while execution
decentralizes, hooks and policies enforce what documentation only
suggests, and subagents do not inherit ambient context.

% ===================================================================
\section{Design Principles}
% ===================================================================

\begin{enumerate}[label=\textbf{P\arabic*}.]
  \item \textbf{Sidecar, not dependency.}  Consumers read files at
    known paths.  No import, no version lock-step, no daemon required.
    Tools adopt ethos at their own pace.
  \item \textbf{Same schema for humans and agents.}  The \field{kind}
    field (\texttt{human} or \texttt{agent}) is the only structural
    difference.  An identity is an identity.
  \item \textbf{Agent definition is a channel binding.}  Like voice
    or email --- the \texttt{.md} file defines tools and workflow,
    ethos defines \emph{who}.
  \item \textbf{Generic extensions, not consumer fields.}  Ethos
    never adds fields for Vox, Beadle, or Biff.  Consumers own their
    own namespace files in the extension directory.
  \item \textbf{Fast.}  Go's $\sim$10\,ms cold start (vs Python's
    $\sim$200\,ms) keeps hook latency invisible.  Ethos is queried on
    session start, on every \cmd{/who}, on every TTS call, and on
    every mission state transition.
  \item \textbf{No daemon.}  All concurrency uses \texttt{flock}
    on lock files.  No background process, no database.
  \item \textbf{Preserve identity content.}  Source \texttt{.md}
    files are the authority on personality, writing style, and talent
    content.  Hooks and persona blocks may restructure but must not
    discard or summarize the underlying meaning.
  \item \textbf{Discipline is a runtime contract, not text.}
    (DES-031.)  Anything that sounds like ``every time,'' ``before,''
    or ``after'' belongs in a hook or a typed primitive, not in a
    personality file.  The mission contract enforces what CLAUDE.md
    can only suggest.
  \item \textbf{Trust-boundary symmetry between read and write paths.}
    (DES-031, DES-036, DES-037.)  Every gate that fires on the write
    path also fires on the read path.  Strict YAML decode
    (\texttt{KnownFields(true)}) applies on \texttt{Save}, on
    \texttt{Load}, and on every sibling-file decode.  An attacker
    with local write access cannot bypass a contract by hand-editing
    a file ethos was about to read.
  \item \textbf{Append-only audit trails are the source of truth;
    derived summaries are convenience.}  (DES-037.)  Memory is
    advisory recall.  Beads track work items.  Mission JSONL event
    logs are the authoritative substrate when something goes wrong:
    one corrupt line does not erase the log, one oversized line does
    not truncate the tail, and the reader degrades line-by-line with
    line-numbered warnings rather than failing whole-file.
\end{enumerate}

% ===================================================================
\section{Package Map}
% ===================================================================

All reusable library code lives under \pkg{internal/}; the public
CLI entry point lives in \pkg{cmd/ethos}.  Module path:
\texttt{github.com/punt-labs/ethos}.

\begin{center}
\begin{longtable}{@{}lp{8.5cm}@{}}
\toprule
\textbf{Package} & \textbf{Responsibility} \\
\midrule
\endfirsthead
\toprule
\textbf{Package} & \textbf{Responsibility} \\
\midrule
\endhead
\pkg{cmd/ethos/}        & CLI entry point: identity, attribute,
                          session, role, team, mission, hook, and
                          admin subcommands; global \cmd{--json} flag \\
\pkg{internal/identity/} & Core identity model (\texttt{Identity}
                          struct), validation, CRUD via \texttt{Store},
                          \texttt{LayeredStore} (repo + global),
                          \texttt{IdentityStore} interface,
                          extension I/O \\
\pkg{internal/attribute/} & Generic CRUD for named markdown files
                          (talents, personalities, writing styles);
                          \texttt{Kind} struct, \texttt{Store} with
                          layered fallback, slug validation \\
\pkg{internal/session/}  & Session roster model (\texttt{Roster},
                          \texttt{Participant}), \texttt{flock}-based
                          concurrent store, PID-keyed session
                          discovery, staleness detection and purge,
                          prefix matching \\
\pkg{internal/process/}  & Process tree walker: find topmost
                          \texttt{claude} ancestor PID via
                          \texttt{/proc} (Linux), \texttt{sysctl}
                          (macOS); result cached via
                          \texttt{sync.Once} \\
\pkg{internal/resolve/}  & 4-step human resolution chain (iam
                          $\to$ git user.name $\to$ git user.email
                          $\to$ \$USER) plus agent resolution from
                          repo config \\
\pkg{internal/hook/}     & Hook handlers (\texttt{SessionStart},
                          \texttt{PreCompact},
                          \texttt{SubagentStart/Stop},
                          \texttt{SessionEnd}, \texttt{FormatOutput});
                          persona builder, team context, extension
                          context, agent file generation, frozen
                          evaluator hash gate (DES-033), verifier
                          context isolation (DES-035) \\
\pkg{internal/mission/}  & Mission contract primitive: typed
                          \texttt{Contract}, flock-protected
                          \texttt{Store}, daily monotonic ID
                          generator, append-only JSONL event log,
                          write-set admission control (DES-032),
                          evaluator hash computation (DES-033),
                          bounded rounds + reflection (DES-034),
                          result artifacts + close gate (DES-036),
                          event log reader (DES-037).  Trust boundary
                          for every Phase 3 invariant. \\
\pkg{internal/mcp/}      & MCP tool definitions and handlers
                          (10~tools via method dispatch);
                          \texttt{Handler} struct with injected
                          identity, session, attribute, role, team,
                          and mission stores \\
\pkg{internal/doctor/}   & Installation health checks: identity
                          directory, human identity, default agent,
                          duplicate fields \\
\pkg{internal/role/}     & Role model (\texttt{Role} struct), CRUD
                          via \texttt{Store}, \texttt{LayeredStore}
                          (repo + global) \\
\pkg{internal/team/}     & Team model (\texttt{Team} struct), CRUD
                          via \texttt{Store}, \texttt{LayeredStore}
                          (repo + global), Z-spec referential
                          integrity validation \\
\bottomrule
\end{longtable}
\end{center}

\subsection{Dependency Graph}

Dependencies flow strictly downward.  No package imports a package
above it in this ordering.

\begin{verbatim}
                       cmd/ethos
       /    /     |    |     |     \    \    \
      /    /      |    |     |      \    \    \
   mcp  hook  mission resolve doctor role team attribute
    | \   | \    |       |
    |  \  |  \   |       |
identity session process |
    |                    |
attribute                |
    +--------------------+
\end{verbatim}

\pkg{cmd/ethos} imports all internal packages directly.
\pkg{internal/mcp} depends on \pkg{identity}, \pkg{session},
\pkg{attribute}, \pkg{role}, \pkg{team}, \pkg{mission},
\pkg{resolve}, \pkg{doctor}, \pkg{hook}, and \pkg{process}.
\pkg{internal/hook} depends on \pkg{identity}, \pkg{session},
\pkg{resolve}, \pkg{process}, \pkg{role}, \pkg{team}, and
\pkg{mission} (the verifier hash gate from DES-033 reads open
missions; the verifier context isolator from DES-035 reads the
matching contract).  \pkg{internal/mission} depends only on
standard library and the YAML decoder; it is a leaf package on
purpose so the trust boundary stays in one place.
\pkg{internal/identity} depends on \pkg{attribute} for slug
validation and content resolution.
\pkg{internal/resolve} depends on \pkg{identity}, \pkg{session},
and \pkg{process}.

% ===================================================================
\section{Identity Model}
% ===================================================================

\subsection{Schema}

An identity is a YAML file at
\fpath{.punt-labs/ethos/identities/<handle>.yaml} (repo-local) or
\fpath{\~{}/.punt-labs/ethos/identities/<handle>.yaml} (global).

\begin{center}
\begin{longtable}{@{}llp{6cm}@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\endfirsthead
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\endhead
\field{name}          & string (required)  & Display name \\
\field{handle}        & string (required)  & Unique lookup key, used as
                                             filename \\
\field{kind}          & string (required)  & \texttt{human} or
                                             \texttt{agent} \\
\field{email}         & string             & Beadle channel binding \\
\field{github}        & string             & Biff channel binding \\
\field{agent}         & string             & Path to Claude Code agent
                                             \texttt{.md} file \\
\field{writing\_style} & string            & Slug referencing a writing
                                             style \texttt{.md} file \\
\field{personality}   & string             & Slug referencing a
                                             personality \texttt{.md}
                                             file \\
\field{talents}       & list of strings    & Slugs referencing talent
                                             \texttt{.md} files \\
\midrule
\multicolumn{3}{@{}l}{\emph{Resolved fields --- populated by
  \texttt{Store.Load}, never persisted:}} \\
\midrule
\field{WritingStyleContent}  & string       & Resolved markdown content
                                              of writing style \\
\field{PersonalityContent}   & string       & Resolved markdown content
                                              of personality \\
\field{TalentContents}       & []string     & Resolved markdown content
                                              per talent (positionally
                                              indexed) \\
\field{Warnings}             & []string     & Warnings from attribute
                                              resolution (missing
                                              \texttt{.md} files) \\
\field{Ext}                  & \texttt{map[string]map[string]string}
                             & Tool-scoped extension data assembled
                               from \fpath{<handle>.ext/} \\
\bottomrule
\end{longtable}
\end{center}

\subsection{Validation}

\texttt{Identity.Validate()} enforces:

\begin{itemize}
  \item \field{name} and \field{handle} are non-empty.
  \item \field{handle} matches
    \verb|^[a-z0-9]([a-z0-9-]*[a-z0-9])?$| (lowercase alphanumeric,
    may contain internal hyphens but not leading/trailing, used as
    filename).
  \item \field{kind} is one of \texttt{human} or \texttt{agent}.
\end{itemize}

Validation errors are returned as a structured
\texttt{ValidationError} type, not strings.

\subsection{Persistence}

\texttt{Store} takes a root directory path and provides:

\begin{center}
\begin{tabular}{@{}lp{7cm}@{}}
\toprule
\textbf{Method} & \textbf{Behavior} \\
\midrule
\texttt{Save(id)}        & Atomic create with \texttt{O\_EXCL} ---
                           rejects duplicate handles.  Validates
                           attribute references.  Creates the
                           extension directory. \\
\texttt{Load(handle)}    & Read YAML, resolve attributes (personality,
                           writing style, talents) to markdown content,
                           assemble extensions from
                           \fpath{<handle>.ext/}. \\
\texttt{List()}          & Return all identities in reference mode
                           (slugs only).  Non-loadable files are
                           reported as warnings. \\
\texttt{FindBy(f, v)}    & Search identities where field \texttt{f}
                           matches value \texttt{v}.  Supported fields:
                           \texttt{handle}, \texttt{email},
                           \texttt{github}. \\
\texttt{Exists(handle)}  & Check file existence. \\
\texttt{Update(h, fn)}   & Read-modify-write with flock.  Applies
                           mutation function \texttt{fn}, validates
                           attribute references, writes atomically
                           via temp file + rename. \\
\texttt{ValidateRefs(id)} & Check that all attribute slugs resolve to
                           existing \texttt{.md} files. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{LayeredStore}

\texttt{LayeredStore} implements \texttt{IdentityStore} with two
layers: repo-local (git-tracked) and user-global.  The repo layer
is optional (nil when not inside a git repository).

\begin{itemize}
  \item \textbf{Load priority}: repo first, then global.  Parse
    errors from the repo layer are surfaced (not silently fallen
    through).  File-not-found falls through to global.
  \item \textbf{Extensions}: always resolve from the global store,
    even when the identity itself is in repo.
  \item \textbf{Attribute resolution}: tries the identity's own
    layer first, falls back to the other layer for any missing
    \texttt{.md} files.
  \item \textbf{List}: returns identities from both stores,
    deduplicated by handle.  Repo wins on collision.
  \item \textbf{Save}: writes to repo if available, otherwise
    global.  Extension directory is always created in global.
  \item \textbf{Update}: updates the identity in the owning store
    (repo if it exists there, otherwise global).  Uses cross-layer
    \texttt{ValidateRefs} so attribute references in either store
    are accepted.
\end{itemize}

\subsection{IdentityStore Interface}

The \texttt{IdentityStore} interface abstracts over both
\texttt{Store} (single-root) and \texttt{LayeredStore} (repo +
global).  All consumers --- MCP handlers, hooks, CLI commands ---
accept \texttt{IdentityStore} rather than a concrete type.  This
also lets the mission package's \texttt{NewLiveHashSources}
(DES-033) consume the same interface as the persona-animation hooks.

% ===================================================================
\section{Attribute System}
% ===================================================================

Attributes are named markdown documents stored as
\fpath{<kind>/<slug>.md} files under both repo-local and global ethos
roots.  Three kinds are pre-defined:

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Kind} & \textbf{Directory} & \textbf{CLI Command} &
  \textbf{Identity Field} \\
\midrule
Talents       & \texttt{talents/}       & \cmd{talent}
              & \field{talents} (list) \\
Personalities & \texttt{personalities/} & \cmd{personality}
              & \field{personality} (scalar) \\
Writing Styles & \texttt{writing-styles/} & \cmd{writing-style}
              & \field{writing\_style} (scalar) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Kind and Store}

The \texttt{Kind} struct configures a \texttt{Store} for one attribute
type, carrying directory name, CLI command name, display name, and
plural name.  \texttt{Store} provides \texttt{Save}, \texttt{Load},
\texttt{List}, \texttt{Delete}, \texttt{Exists}, and \texttt{Path}.

\texttt{Store} supports an optional fallback store.  When set,
read operations check the fallback (repo layer) first, then the
primary store (global layer).  Writes always go to the primary store.

\subsection{Slug Validation}

Slugs follow the same pattern as identity handles:
\verb|^[a-z0-9]([a-z0-9-]*[a-z0-9])?$|.  Path traversal is
prevented by resolving the candidate path and checking it stays
within the attribute directory.

% ===================================================================
\section{Extension System}
% ===================================================================

Consumers of ethos (Beadle, Biff, Vox) store tool-specific attributes
on identities without modifying the ethos schema.  Extensions use a
namespaced, file-per-tool design.

\subsection{Storage Layout}

\begin{verbatim}
~/.punt-labs/ethos/identities/
  mal.yaml                     # ethos owns -- core identity fields
  mal.ext/                     # extension directory
    beadle.yaml                # beadle owns -- GPG key, IMAP config
    biff.yaml                  # biff owns -- preferred TTY
    vox.yaml                   # vox owns -- default mood
\end{verbatim}

Each file is a flat YAML map of key-value pairs.  Ethos never reads
or interprets extension contents except to assemble the merged view
returned by \texttt{Load()} and the \tool{identity} MCP tool.

\subsection{Ownership Rules}

\begin{enumerate}
  \item Ethos creates the \fpath{<handle>.ext/} directory on identity
    creation.  Extensions always live in the global store, even when
    the identity is repo-local.
  \item Each tool manages its own \fpath{<namespace>.yaml} file.
  \item Tools may read/write their file directly (sidecar contract) or
    through the ethos CLI/MCP interface.
  \item Tools must not write to another tool's namespace file.
\end{enumerate}

\subsection{Validation Constraints}

\begin{center}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Field} & \textbf{Pattern} & \textbf{Limit} \\
\midrule
Namespace   & \verb|^[a-z][a-z0-9-]*$|    & 32 chars \\
Key         & \verb|^[a-z][a-z0-9_]*$|    & 64 chars \\
Value       & Any YAML scalar             & 4096 bytes \\
Keys per namespace    & ---               & 64 \\
Namespaces per persona & ---              & 32 \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Two Scopes}

Extensions exist at two independent scopes with different types:

\begin{enumerate}
  \item \textbf{Persona-level} (durable) --- files in
    \fpath{<handle>.ext/}.  Flat \texttt{map[string]string} (YAML
    scalars only).  Persist across sessions.
  \item \textbf{Session-participant-level} (ephemeral) --- the
    \field{ext} field on each \texttt{Participant} in the session
    roster.  Typed as \texttt{map[string]any} (arbitrary nested
    values).  Deleted when the session ends.
\end{enumerate}

\subsection{Session Context (DES-022)}

Any extension namespace may include a \texttt{session\_context} key.
During \texttt{SessionStart}, \texttt{PreCompact}, and
\texttt{SubagentStart} hooks, ethos collects all
\texttt{session\_context} values from the agent's persona-level
extensions, joins them, and emits the result alongside the persona
block.  Consumers (Biff, Quarry, Vox) inject behavioral instructions
into the model's context without ethos understanding the content.

% ===================================================================
\section{Session Roster}
% ===================================================================

The session roster provides multi-participant awareness for Claude
Code sessions.  It answers two questions: ``Who am I?'' and ``Who is
everyone else?''

\subsection{Data Model}

Stored at \fpath{\~{}/.punt-labs/ethos/sessions/<session-id>.yaml}:

\begin{center}
\begin{longtable}{@{}llp{5cm}@{}}
\toprule
\textbf{Field} & \textbf{Source} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Field} & \textbf{Source} & \textbf{Purpose} \\
\midrule
\endhead
\field{session}      & Hook input        & Session identifier \\
\field{started}      & Roster creation   & RFC~3339 timestamp \\
\field{repo}         & Git remote        & Repository name
                                           (org/repo) \\
\field{host}         & \texttt{os.Hostname()} & Short hostname \\
\field{participants} & Hook lifecycle    & Ordered list of
                                           \texttt{Participant}
                                           records \\
\bottomrule
\end{longtable}
\end{center}

Each participant record carries \field{agent\_id}, \field{persona},
\field{agent\_type}, \field{parent}, \field{joined}, and a
tool-scoped ephemeral \field{ext} map.

\subsection{Tree Structure}

Participants form a tree.  The root (no parent) is the human user.
The primary Claude agent is the root's child.  Subagents are children
of the primary agent.

\begin{verbatim}
mal (root -- $USER)
  +-- 19147 (archie -- Claude PID)
       +-- a5734dd (code-reviewer)
       +-- a93c2be (silent-failure-hunter)
       +-- b8823ff (Explore, no persona)
\end{verbatim}

Any participant can derive relationships from the tree: root (walk up
to \field{parent:~empty}), initiator (my \field{parent}), delegates
(anyone whose \field{parent} is my \field{agent\_id}), siblings (same
\field{parent}).

\subsection{Lifecycle via Hooks}

\begin{center}
\begin{tabular}{@{}llp{4.5cm}@{}}
\toprule
\textbf{Hook Event} & \textbf{Roster Operation} & \textbf{Data} \\
\midrule
\texttt{SessionStart}  & Create roster, join root + primary; emit
                          persona block, team context, working
                          context (Phase 4.1); generate agent files
                        & \field{session\_id}, \texttt{\$USER},
                          Claude PID \\
\texttt{SubagentStart}  & Join subagent; gate verifier spawns
                          (DES-033, DES-035); emit persona OR
                          verifier isolation block
                        & \field{session\_id}, \field{agent\_id},
                          \field{agent\_type} \\
\texttt{SubagentStop}   & Leave subagent
                        & \field{session\_id}, \field{agent\_id} \\
\texttt{PreCompact}     & Re-emit full persona block, team context,
                          extension context
                        & \field{session\_id} \\
\texttt{PostToolUse}    & Format MCP tool output; audit log (Phase
                          4.3)
                        & tool name, output \\
\texttt{SessionEnd}     & Delete roster + PID-keyed session file
                        & \field{session\_id} \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Concurrency}

Multiple \texttt{SubagentStart} hooks may fire in parallel.  Roster
writes use \texttt{flock(LOCK\_EX)} on a per-session lock file
(\fpath{sessions/<session-id>.lock}) before read-modify-write.  Roster
files are written atomically via temp file + rename.

\subsection{Session Discovery}

Non-hook callers (Biff, Vox) need the session ID without receiving it
on stdin.  The \texttt{SessionStart} hook writes the session ID to a
PID-keyed file:

\begin{verbatim}
~/.punt-labs/ethos/sessions/current/<claude-pid>
\end{verbatim}

Any descendant process walks the process tree to the topmost
\texttt{claude} ancestor PID, reads that file, and gets the session
ID.  The \texttt{SessionEnd} hook deletes the file.

\subsection{Staleness Detection}

\texttt{Store.Purge()} cleans up rosters from crashed sessions:

\begin{enumerate}
  \item Extract the primary agent's \field{agent\_id} (second
    participant).
  \item If numeric: check whether the PID is alive via
    \texttt{os.FindProcess} + signal~0.  If dead, treat as stale
    immediately.
  \item If non-numeric: age-based fallback (24-hour TTL).
  \item Rosters with fewer than 2~participants are always stale.
\end{enumerate}

\texttt{PurgeCurrent()} additionally cleans stale PID-keyed files
from the \fpath{sessions/current/} directory when the referenced
process is no longer alive.

% ===================================================================
\section{Teams and Roles}
% ===================================================================

\subsection{Role Model}

A role is a named YAML file at \fpath{roles/<slug>.yaml}.

\begin{center}
\begin{tabular}{@{}llp{6.5cm}@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\field{name}             & string       & Slug-format name \\
\field{model}            & string (optional) & Claude model preference;
                                          one of \texttt{opus},
                                          \texttt{sonnet}, \texttt{haiku},
                                          \texttt{inherit}, or a full
                                          \texttt{claude-*} ID. \\
\field{responsibilities} & []string     & What the role does \\
\field{permissions}      & []string     & What the role may access \\
\field{tools}            & []string     & MCP tools available to agents
                                          in this role (used in agent
                                          file generation) \\
\field{anti\_responsibilities} & []string & What the role does NOT do
                                          (Phase 2.1; derived from the
                                          team collaboration graph
                                          when omitted) \\
\field{output\_format}   & string (optional) & Markdown template for
                                          structured handoff (Phase
                                          2.3) \\
\field{safety\_constraints} & []string  & Pre-tool denylist
                                          patterns (Phase 4.2) \\
\bottomrule
\end{tabular}
\end{center}

Role names follow slug validation.  CRUD is provided by
\texttt{role.Store}, with \texttt{role.LayeredStore} searching both
repo-local and global roots.  Deletion is guarded by referential
integrity: a role cannot be deleted if any team references it.

\subsection{Team Model}

A team is a named YAML file at \fpath{teams/<slug>.yaml}.

\begin{center}
\begin{tabular}{@{}llp{5.5cm}@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\field{name}           & string           & Slug-format name \\
\field{repositories}   & []string         & Repos this team owns
                                            (org/repo format) \\
\field{members}        & []Member         & Identity-role
                                            bindings \\
\field{collaborations} & []Collaboration  & Directed relationships
                                            between roles \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Member}: \field{identity} (handle) + \field{role} (slug).

\textbf{Collaboration}: \field{from} (role) $\to$ \field{to} (role),
with \field{type} constrained to \texttt{reports\_to},
\texttt{collaborates\_with}, or \texttt{delegates\_to}.

\subsection{Z-Spec Invariants}

\texttt{team.Validate()} enforces referential integrity via
callback functions (no circular imports):

\begin{itemize}
  \item Team name follows slug rules.
  \item At least one member.
  \item Every member's identity exists.
  \item Every member's role exists.
  \item No duplicate identity/role pairs.
  \item Every collaboration's \field{from} and \field{to} roles are
    filled by at least one member.
  \item No self-collaborations.
  \item Collaboration \field{type} is one of the three valid values.
\end{itemize}

These invariants are checked on team creation and member addition.
Phase 3.5 (DES-035) adds an additional cross-cutting check: a
mission's worker and evaluator handles cannot share a team-scoped
role binding under canonicalization.

% ===================================================================
\section{Identity Resolution}
% ===================================================================

\subsection{Human Resolution}

\texttt{resolve.Resolve(store, sessionStore)} determines the active
identity handle through a 4-step priority chain.  The chain stops
at the first match:

\begin{enumerate}
  \item \textbf{iam declaration} --- walk the process tree via
    \texttt{FindClaudePID()}, read the PID-keyed session file, load
    the roster, find the matching participant.  If found with a
    persona, return it.  If found with no persona, stop (do not fall
    through --- this is an explicit ``no identity'').
  \item \textbf{git config user.name} --- match against identity
    \field{github} field.
  \item \textbf{git config user.email} --- match against identity
    \field{email} field.
  \item \textbf{\$USER} --- match against identity \field{handle}
    field.
\end{enumerate}

Returns an error when no step matches.

\subsection{Agent Resolution}

\texttt{resolve.ResolveAgent(repoRoot) (string, error)} reads the
repo-local configuration file at \fpath{.punt-labs/ethos.yaml}.
The \texttt{RepoConfig} struct also carries a \field{team} field,
used by \texttt{resolve.ResolveTeam(repoRoot) (string, error)} to
determine which team owns the current repository.  Both functions
return \texttt{("", nil)} when the repository is not configured or
when \texttt{repoRoot} is empty; non-nil errors propagate YAML
parse failures and permission errors from the underlying
\texttt{LoadRepoConfig} call with a \texttt{"resolve agent"} or
\texttt{"resolve team"} wrap prefix.  Callers decide fail-open vs
fail-closed per operational role: \texttt{HandleSessionStart}
propagates the error up (fail-closed at the binary level, with
the shell wrapper's \texttt{|| true} preserving Claude Code
session fail-open semantics), \texttt{BuildTeamSection} in the
pre-compact hook logs and returns an empty string (fail-open per
its documented contract), \cmd{ethos resolve-agent} exits
non-zero, and \texttt{CheckDefaultAgent} in the doctor package
surfaces the error as a diagnostic state (ethos-dc0).

% ===================================================================
\section{Process Tree Walker}
% ===================================================================

\texttt{process.FindClaudePID()} identifies the Claude Code instance
that spawned the current process.  It is used to determine the
primary agent's \field{agent\_id} for session rosters and to locate
PID-keyed session files.

\begin{enumerate}
  \item Read the process table using native OS interfaces:
    \texttt{/proc/<pid>/stat} on Linux,
    \texttt{sysctl kern.proc.pid} on macOS.  No subprocess spawning.
  \item Walk from the current PID upward, following parent pointers.
  \item Track the topmost PID where the command name is
    \texttt{claude}.
  \item Cap the walk at 10~iterations to prevent infinite loops in
    degenerate process trees.
  \item If no \texttt{claude} ancestor is found, fall back to
    \texttt{os.Getppid()}.
  \item Cache the result for the lifetime of the process via
    \texttt{sync.Once}.
\end{enumerate}

\subsection{Version-Named Binary Handling}

Claude Code installs version-named binaries under
\fpath{\~{}/.local/share/claude/versions/<version>}.  The process
command name for these binaries is the version string (e.g.,
\texttt{2.1.86}), not \texttt{claude}.
\texttt{normalizeClaudeComm(comm, exePath)} detects this pattern:
if the executable path contains \texttt{/claude/versions/}, the
command name is normalized to \texttt{claude}.  On Linux, the exe
path is read from \texttt{/proc/<pid>/exe}.  On macOS, it is read
from \texttt{kern.procargs2}.

% ===================================================================
\section{Persona Animation}
% ===================================================================

Persona animation is the process of injecting identity context into
Claude Code sessions so the model behaves according to the defined
personality, writing style, team role, and (per DES-031) the active
mission contract.

\subsection{Four Layers}

The complete agent context has four layers, each with a different
lifecycle:

\begin{enumerate}
  \item \textbf{Persona} (durable; identity) --- personality, writing
    style, talents.  Survives across tasks and sessions.  Pinned
    via DES-033 hash for evaluators.
  \item \textbf{Role} (semi-durable; team) --- tools,
    responsibilities, anti-responsibilities, team position.
    Changes when the team changes.
  \item \textbf{Mission} (ephemeral; per-task) --- typed I/O contract,
    write-set, success criteria, evaluator, budget, and the
    appended event log.  One delegation, then closed.  See
    \S\ref{sec:mission}.
  \item \textbf{Session} (ephemeral; per-process) --- the participant
    roster, working context (Phase 4.1: git branch, uncommitted
    changes, recent issues), and ephemeral session-participant ext.
\end{enumerate}

The first three are emitted by hooks; the fourth is derived from
the runtime.

\subsection{Hook Phases}

\begin{enumerate}
  \item \textbf{SessionStart} --- emits the full persona block for
    the repo's default agent.  Includes personality (with opening
    line fold), writing style, talents, extension context
    (\texttt{session\_context} keys), team context (team members,
    their roles, personalities, collaborations), and (Phase 4.1)
    working context.  Also generates
    \fpath{.claude/agents/<handle>.md} files from identity data and
    installs static agent definitions from the ethos agents
    directory.

  \item \textbf{PreCompact} --- fires before context compaction.
    Re-emits the full persona block, extension context, and team
    context so behavioral instructions survive compaction.

  \item \textbf{SubagentStart} --- auto-matches the subagent's
    \field{agent\_type} against ethos identities.  Three branches:
    \begin{itemize}
      \item \emph{Verifier spawn (DES-033, DES-035)}: the hook
        detects that the spawning identity is the
        \field{Evaluator.Handle} of one or more open missions.  It
        recomputes the evaluator content hash and compares against
        every matching pinned hash.  On mismatch the spawn is
        REFUSED with a multi-line drift error naming every drifted
        mission.  On match it REPLACES the persona block with a
        verifier isolation block (mission contract bytes,
        verification criteria, file allowlist) and a MUST-NOT
        directive against parent transcript and worker scratch.
      \item \emph{Persona spawn}: emits the subagent's persona
        block, extension context, and a ``You report to'' line
        referencing the parent agent from the session roster.
      \item \emph{Unmatched spawn}: no injection; the subagent
        operates with its definition file alone.
    \end{itemize}
\end{enumerate}

\subsection{Persona Block Construction}

\texttt{BuildPersonaBlock(id)} assembles the persona from a resolved
identity:

\begin{itemize}
  \item Opening line: ``You are Name (handle), <first prose sentence
    of personality>.''
  \item Remaining personality content (first paragraph skipped to
    avoid redundancy).
  \item Writing style under \texttt{\#\# Writing Style}.
  \item Talent slugs under \texttt{\#\# Talents} (slug list, not
    full content, to stay within context budget).
\end{itemize}

Leading top-level headings (\texttt{\# Title}) are stripped from
personality and writing style content.  Sub-headings are preserved.

\subsection{Team Context Construction}

\texttt{BuildTeamContext(team, roles, identities, selfHandle)} emits
a team section with per-member entries.  For teammates (not self),
each entry includes full personality content, writing style, role
responsibilities, and talent slugs.  The self entry is compact (role
and responsibilities only) since the full persona block was already
emitted.  Collaborations are listed as compact directed edges.

\subsection{Constraints}

\begin{itemize}
  \item Persona context is additive --- it does not override CLAUDE.md.
    Project-level instructions win on conflicts.
  \item Agent definition format is unchanged --- \fpath{.claude/agents/*.md}
    files remain standard Claude Code agent definitions.
  \item Ethos is not required --- if not installed, no persona is injected.
    Agents work normally.
  \item Talent content is not injected inline --- talents are listed as
    slugs.  Full content is available on demand via MCP.
  \item Verifier spawns receive ONLY the mission contract,
    verification criteria, and file allowlist --- the persona block
    is suppressed for evaluator subagents to enforce DES-035's
    independence requirement.
\end{itemize}

% ===================================================================
\section{Mission Primitive}
\label{sec:mission}
% ===================================================================

The mission primitive is the workflow half of ethos.  Where the
identity layer answers ``who is the agent,'' the mission contract
answers ``what is this delegation, what is the worker allowed to
touch, and how do we verify the result.''  Phase 3 (DES-031 through
DES-037) ships seven primitives that compose into a single typed
runtime contract enforced at the store boundary.

\subsection{Why a Contract}

Free-form delegation between agents fails three concrete ways: the
leader changes the success criteria mid-cycle and the worker
implements to a moving target; the reviewer at round 3 is a
different agent than the reviewer at round 1, with a different
threshold; and two workers edit the same file concurrently because
no mechanism declared the conflict.  These failures map onto the
four architecture rules from
\texttt{\~{}/Documents/agents-architecture.tex} (roles are
interfaces, centralize understanding, hooks enforce, subagents do
not inherit ambient context) --- but the rules are unenforceable as
prose in CLAUDE.md.  The mission contract turns each rule into a
runtime check at the \pkg{internal/mission} store boundary.

\subsection{Schema}

A mission contract is a YAML file at
\fpath{\~{}/.punt-labs/ethos/missions/<id>.yaml}.

\begin{center}
\begin{longtable}{@{}llp{7cm}@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\endfirsthead
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\endhead
\field{mission\_id}    & string         & \verb|m-YYYY-MM-DD-NNN|.
                                          Server-controlled. \\
\field{status}         & string         & \texttt{open} on create;
                                          terminal states are
                                          \texttt{closed},
                                          \texttt{failed},
                                          \texttt{escalated}. \\
\field{created\_at}    & RFC3339        & Server-controlled. \\
\field{updated\_at}    & RFC3339        & Server-controlled. \\
\field{closed\_at}     & RFC3339        & Server-controlled on
                                          terminal transition. \\
\field{leader}         & handle         & Who launched the mission. \\
\field{worker}         & handle         & Who executes. \\
\field{evaluator}      & object         & \field{handle},
                                          \field{pinned\_at},
                                          \field{hash}.  Server
                                          fills timestamp and hash. \\
\field{inputs}         & object         & Bead ID, files the worker
                                          MUST read, references the
                                          worker MAY read. \\
\field{write\_set}     & []string       & File allowlist; non-empty;
                                          per-entry validated for
                                          traversal, absolute paths,
                                          control bytes; segment
                                          normalization for conflict
                                          detection. \\
\field{tools}          & []string       & Optional tool allowlist. \\
\field{success\_criteria} & []string    & Non-empty; free-form
                                          strings the evaluator
                                          interprets. \\
\field{budget}         & object         & \field{rounds} (1--10,
                                          default 3) and
                                          \field{reflection\_after\_each}
                                          (bool). \\
\field{current\_round} & int            & 1-indexed; advances via
                                          \cmd{mission advance}
                                          gate (DES-034). \\
\field{context}        & string         & Optional prose context. \\
\bottomrule
\end{longtable}
\end{center}

\subsection{Storage Layout}

\begin{verbatim}
~/.punt-labs/ethos/missions/
  m-2026-04-08-005.yaml          # contract (Phase 3.1)
  m-2026-04-08-005.jsonl         # append-only event log (3.1)
  m-2026-04-08-005.lock          # per-mission flock target (3.1)
  m-2026-04-08-005.reflections.yaml  # round-by-round reflections (3.4)
  m-2026-04-08-005.results.yaml      # round-by-round results (3.6)
  .counter-2026-04-08            # daily ID counter, flock-protected
  .create.lock                   # directory create-lock (3.2)
\end{verbatim}

The contract, the JSONL event log, and the two sibling YAML files
all live under the same per-mission flock at \fpath{<id>.lock}.
Contract+reflection and contract+result operations are atomic with
respect to concurrent readers.  The directory-level
\fpath{.create.lock} is a separate stable inode acquired by
\texttt{Store.Create} for the duration of the conflict scan and the
new mission's write --- two concurrent creates with disjoint mission
IDs cannot race past the conflict scan (DES-032).

\subsection{Lifecycle}

\begin{enumerate}
  \item \textbf{Create.}  The leader writes a complete YAML file
    and runs \cmd{ethos mission create -f mission.yaml} (or calls
    the MCP \tool{mission} tool with \field{method} \texttt{create}).
    The store decodes strictly (\texttt{KnownFields(true)}, no
    unknown fields, no multi-document YAML, no trailing content),
    runs \texttt{Validate()}, fills server-controlled fields, runs
    the write-set conflict scan against every open mission, runs
    the role-overlap check against worker and evaluator, computes
    the evaluator content hash from the live identity store, and
    appends a \texttt{create} event to the JSONL log.  Any
    rejection burns a daily counter slot but commits no other
    state.
  \item \textbf{Round 1.}  The worker executes against the contract.
    \texttt{CurrentRound} starts at 1; no advance gate fires for
    the first round.
  \item \textbf{Reflect.}  At the end of each round the leader runs
    \cmd{ethos mission reflect -f reflection.yaml} to submit a
    typed reflection (\field{round}, \field{converging},
    \field{signals}, \field{recommendation},
    optional \field{reason}).  Reflections are append-only;
    duplicates and out-of-order rounds are refused.  A
    \texttt{reflect} event lands in the JSONL log.
  \item \textbf{Advance.}  \cmd{ethos mission advance} bumps
    \field{current\_round}.  The advance gate refuses if the
    current round has no reflection on file, if the reflection's
    recommendation is terminal (\texttt{stop} or \texttt{escalate}),
    or if \field{current\_round} would exceed \field{budget.rounds}.
    The order of these checks is load-bearing: the operator sees
    ``close and re-scope'' on budget exhaustion, not ``submit one
    more reflection'' on the last round.  A
    \texttt{round\_advanced} event lands in the JSONL log.
  \item \textbf{Result.}  At any point in the current round, the
    worker submits a typed result via \cmd{ethos mission result -f
    result.yaml} (\field{verdict}, \field{confidence},
    \field{files\_changed}, \field{evidence},
    \field{open\_questions}, \field{prose}).  Results are
    append-only and per-round unique.  Each result's
    \field{files\_changed} paths are validated against the
    contract's write-set via the asymmetric
    \texttt{pathContainedBy} helper (DES-036).  A \texttt{result}
    event lands in the JSONL log.
  \item \textbf{Close.}  \cmd{ethos mission close} transitions to
    a terminal status.  The close gate (DES-036) refuses every
    terminal transition without a valid result artifact for the
    current round.  The refusal message names the mission, the
    round, and the submission command.  No override flag exists.
    A \texttt{close} event lands in the JSONL log carrying the
    satisfying result's round and verdict.
\end{enumerate}

\paragraph{Verified Lifecycle.}
The mission lifecycle is not merely specified --- it is tested at four
levels of the test pyramid.  L2 subprocess tests verify that each CLI
command (\cmd{create}, \cmd{reflect}, \cmd{advance}, \cmd{result},
\cmd{close}) exits with the correct code and produces the expected
output.  L2.5 in-process handler tests (DES-042) exercise every gate
--- the write-set conflict scan, the evaluator hash gate (DES-033),
the round-advance gate (DES-034), and the close gate (DES-036) ---
through direct \texttt{RunE} invocation with captured stdout/stderr.
L3 MCP integration tests drive the same gates through the
\texttt{mission} MCP tool via a real stdio client.  L4 behavioral
tests (DES-043) spawn real Claude Code agents against mission
contracts and verify protocol compliance end-to-end: the agent reads
the contract before writing, submits a typed result artifact, and
stays within the declared write-set even under adversarial prompting.
Together, these layers verify the entire path from typed contract to
terminal close event.

\subsection{Write-Set Admission Control (DES-032)}

\texttt{Store.Create} rejects a new mission whose \field{write\_set}
overlaps any currently-open mission's \field{write\_set}.  The check
runs at create time, not at first edit, so the conflict surfaces
before the worker runs.

\textbf{Conflict semantics: segment-prefix overlap on cleaned paths.}
Two paths overlap when, after normalization (trim whitespace, replace
backslashes, trim trailing slash, split on \texttt{/}, drop empty
segments, drop \texttt{.} segments), one path's segment list is a
prefix of the other's.  \texttt{internal/foo} overlaps
\texttt{internal/foo/bar.go} (forward prefix); \texttt{internal/foo}
does NOT overlap \texttt{internal/foobar} (segment boundary).
Comparison is case-sensitive (POSIX).  Round 3 added empty-segment
filtering (\texttt{internal//foo} normalizes to
\texttt{[internal foo]}); round 4 added single-dot filtering
(\texttt{./internal/foo} normalizes to \texttt{[internal foo]}) ---
the equivalence-class fix that the frozen evaluator caught and the
leader's earlier review missed.

\textbf{Active mission set.}  Only \texttt{Status == StatusOpen}
missions enter the conflict scan.  Closing a mission is the explicit
way to release its write-set claim.

\textbf{Two-level locking.}  The per-mission flock from DES-031
serializes writes to a single mission.  DES-032 adds a directory-
level \fpath{.create.lock} acquired by \texttt{Store.Create} for the
conflict scan AND the new mission's write.  Without the directory
lock, two concurrent creates with disjoint mission IDs would each
acquire their own per-mission lock, both pass the conflict scan, and
both write --- a TOCTOU race the per-mission lock cannot close.

\textbf{Fatal load failure.}  If \texttt{Store.Load} fails on any
existing mission during the scan (file corrupted, permission denied,
hand-edited beyond strict YAML), the entire create call fails with a
wrapped error naming the unloadable mission.  Silently skipping a
corrupt mission would defeat the gate --- an attacker could bypass
the check by breaking exactly the existing mission whose write-set
blocks them.

\subsection{Frozen Evaluator (DES-033)}

The mission contract names its evaluator at launch.  At create time,
\texttt{Store.ApplyServerFields} resolves the evaluator handle to its
full identity content (personality, writing\_style, talents, role),
computes a deterministic SHA-256 over every content source, and
writes the hex-encoded result into \field{Evaluator.Hash}.  An
unresolvable evaluator handle fails the create --- an empty hash is
not a valid outcome.

At verifier subagent spawn time, the \texttt{SubagentStart} hook
recomputes the hash from current identity content and compares
against every open mission whose \field{Evaluator.Handle} matches
the spawning subagent.  Any mismatch is a fatal refusal.  The
mismatch error names every drifted mission, the pinned and current
hash prefixes, a per-section breakdown of the current content, and
both recovery paths (revert the edit, or close and relaunch).

\textbf{Hash algorithm} (\texttt{ethos-evaluator-hash-v1}, version-
prefixed for future format changes):

\begin{itemize}
  \item Format version included in the hash input so a bump
    invalidates every prior pinned hash.
  \item Field separators 0x1F (Unit Separator) between label and
    value, 0x1E (Record Separator) between sections.  Both are ASCII
    control bytes that ethos identity validation already rejects from
    handles and slugs.
  \item Length prefixes in BYTES, not runes, to prevent field-
    boundary attacks.
  \item Section order: handle, personality content, writing style
    content, talents (declaration order), roles (lexicographic by
    \texttt{team/role\_name}).
  \item Canonical role content rendered as \texttt{key=value$\backslash$n},
    NOT \texttt{yaml.Marshal} (which is nondeterministic across
    releases).
\end{itemize}

\textbf{Role binding semantics.}  Ethos identities have no direct
role field --- roles bind via team membership.
\texttt{NewLiveHashSources} walks every team, collects every
\texttt{(team, role)} assignment for the evaluator handle, and
includes each with the team name as a prefix so two identical role
names on different teams stay distinguishable.

\textbf{Verifier hook enforcement.}

\begin{itemize}
  \item Runs BEFORE joining the session roster.  A refused spawn
    leaves no roster trace.
  \item Computes the current hash ONCE per hook invocation (cached
    across all matching missions) and aggregates every mismatch into
    a single multi-line error.
  \item A corrupt or unloadable mission is fatal, not silently
    skipped.  Misconfiguration (non-nil mission store, nil
    \texttt{HashSources}) is fatal.  Silent skip would let stale
    evaluator content through under a configuration error.
  \item No runtime bypass.  No flag, no environment variable, no
    configuration option to disable the gate.  The only ways to
    proceed after a mismatch are: revert the edit, close the
    mission, or fix the misconfigured installation.
\end{itemize}

\subsection{Bounded Rounds with Reflection (DES-034)}

A typed \texttt{Reflection} artifact sits between round N and round
N+1.  The round-advance gate in \texttt{Store.AdvanceRound} refuses
to begin round N+1 until the reflection for round N is on disk AND
its recommendation is non-terminal AND the budget isn't exhausted.

\textbf{Reflection schema} (\fpath{internal/mission/reflection.go}):

\begin{itemize}
  \item \field{round} --- integer, the round just ended
  \item \field{created\_at} --- RFC3339 server-filled at append time
  \item \field{author} --- leader handle recording the reflection
  \item \field{converging} --- bool
  \item \field{signals} --- list of observations (at least one,
    each non-empty, no control characters)
  \item \field{recommendation} --- enum:
    \texttt{continue} $|$ \texttt{pivot} $|$ \texttt{stop} $|$
    \texttt{escalate}
  \item \field{reason} --- prose, required when recommendation is
    terminal
\end{itemize}

\textbf{Recommendation semantics.}

\begin{itemize}
  \item \texttt{continue} --- advance permitted, same approach
  \item \texttt{pivot} --- advance permitted, worker takes a
    different approach in round N+1
  \item \texttt{stop} --- advance refused, mission must close
  \item \texttt{escalate} --- advance refused, mission must be
    re-scoped or escalated to a human
\end{itemize}

\textbf{Sibling-file storage.}  Reflections live in
\fpath{<id>.reflections.yaml}, NOT inside the contract.  Two
reasons: the contract is pinned at launch per DES-031 and an
unbounded reflection slice inside it would force every
\texttt{Store.Update} to rewrite an unbounded history; and the
reflections file grows as rounds happen, while the contract file
stays structurally stable.

\textbf{Append-only invariant.}  \texttt{AppendReflection} refuses
to overwrite an existing round's reflection.  The reflections file
is monotone-sorted by round number at decode time; a hand-edited or
corrupt out-of-order file is rejected.

\textbf{Round tracking.}  \field{Contract.CurrentRound} is a small
integer cached on the contract.  Default-filled to 1 on Create and
Load so pre-3.4 contracts upgrade cleanly.  Validation enforces
\texttt{CurrentRound in [1, Budget.Rounds]}.  Phase 3.7's event log
reader can reconstruct the round history from events for audit ---
the field is a cache of derivable state, not a source of truth.

\subsection{Verifier Isolation (DES-035)}

Two independent runtime gates enforce verifier independence from the
implementer:

\textbf{1. Role-overlap check at \texttt{Store.Create}.}  A
\texttt{RoleLister} interface and \texttt{Store.WithRoleLister(r)}
opt-in method.  When set, \texttt{Store.Create} refuses a contract
whose worker and evaluator handles either match literally or share
a team-scoped role binding under canonicalization (substring after
the last \texttt{/}).  Distinct roles on the same team are accepted
--- the canonical ethos pattern is an implementer and a security
reviewer on the same team verifying each other's work.

\textbf{2. Verifier context isolation at \texttt{SubagentStart}.}
When the hook detects a verifier spawn (the existing DES-033
\texttt{checkVerifierHash} discriminator), the hook REPLACES the
subagent's \texttt{additionalContext} with a structured isolation
block instead of the normal persona block.  The block contains:

\begin{itemize}
  \item \texttt{\#\# Verifier context (mission <id>)}
  \item \texttt{\#\#\# Mission contract} --- byte-for-byte from disk
    via \texttt{Store.ContractPath} + \texttt{os.ReadFile}, NOT
    re-marshaled (which would reorder keys or drop YAML comments)
  \item \texttt{\#\#\# Verification criteria} --- the contract's
    \field{success\_criteria} list, verbatim
  \item \texttt{\#\#\# File allowlist} --- two labeled sub-sections:
    repo-relative paths and absolute paths
  \item A \texttt{MUST NOT} directive: the verifier may not read
    parent transcript, worker scratch, or paths outside the
    allowlist
\end{itemize}

The persona block is EXCLUDED for verifier spawns.  Parent
transcript is excluded by structural replacement (the hook's
\texttt{additionalContext} field is overwritten, not appended).
For multi-mission evaluators, the block concatenates one section
per mission separated by \texttt{---}.

\subsection{Result Artifacts and Close Gate (DES-036)}

A typed \texttt{mission.Result} artifact pins a closed schema with
strict YAML decoding, full validation, and append-only sibling
storage at \fpath{<id>.results.yaml}.

\textbf{Schema.}

\begin{itemize}
  \item \field{mission} --- mission ID; must equal the target file's
    mission (DES-036 round 5 trust-symmetry fix)
  \item \field{round} --- integer; must equal the contract's current
    round
  \item \field{author} --- worker handle
  \item \field{verdict} --- enum: \texttt{pass} $|$ \texttt{fail}
    $|$ \texttt{escalate}
  \item \field{confidence} --- float in $[0.0, 1.0]$ excluding NaN
  \item \field{files\_changed} --- list of \texttt{(path, added,
    removed)} tuples; every path validated against the contract's
    write-set via the asymmetric \texttt{pathContainedBy} helper
  \item \field{evidence} --- non-empty list of \texttt{(name,
    status)} tuples
  \item \field{open\_questions} --- optional list of strings
  \item \field{prose} --- optional human-facing summary; accepts
    \texttt{$\backslash$n$\backslash$r$\backslash$t} only
\end{itemize}

\textbf{Asymmetric containment.}  \texttt{pathContainedBy(file,
entry)} requires the entry's segment list to be a prefix of the
file's, AND the file to have at least as many segments --- the only
correct shape for ``the file lives inside the allowlist entry.''
The earlier symmetric \texttt{pathsOverlap} (correct for DES-032's
cross-mission conflict check) was caught by all four reviewers as a
parent-prefix exploit and replaced with the new helper.

\textbf{Close gate.}  \texttt{Store.Close} refuses every terminal
transition (\texttt{closed}, \texttt{failed}, \texttt{escalated})
unless \texttt{LoadResults} returns a valid artifact for the
mission's current round.  The refusal message names the mission,
the round, and the submission command.  The gate lives at the store
boundary so CLI and MCP fire it identically --- no override flag.

\textbf{Show payload.}  \texttt{mission.ShowPayload} (in
\fpath{internal/mission/mission.go}) embeds \texttt{*Contract} so
the show JSON shape auto-propagates any future Contract field to
both CLI and MCP without hand-rolling field lists.  Plus
\field{Results} (always an array, never \texttt{null}) and
\field{Warnings} (omitempty) so a \texttt{LoadResults} error
surfaces as structured signal instead of being silently dropped.

\subsection{Append-Only Event Log Reader (DES-037)}

The mission package writes a JSONL event log per mission via a
private \texttt{appendEvent} helper from DES-031.  Every state
transition lands as one line: \texttt{create}, \texttt{update},
\texttt{result}, \texttt{reflect}, \texttt{round\_advanced},
\texttt{verify}, \texttt{close}.  The writer is frozen ---
DES-031 round 3 unexported \texttt{AppendEvent} as a deadlock
footgun, and DES-037 does not re-introduce one.  DES-037 ships the
\emph{reader}.

\textbf{Public API.}
\texttt{Store.LoadEvents(missionID) ([]Event, []string, error)} ---
events in on-disk order, warnings naming any unparseable line
numbers, error for unrecoverable I/O failures.  The 3-tuple shape
departs from \texttt{LoadResults} and \texttt{LoadReflections}
because JSONL degrades per-line by construction --- a single
corrupt line must not fail the whole read.

\textbf{Design pressure: post-mortem first.}  Show as much of what
happened as possible, even if the file is partially damaged.

\begin{itemize}
  \item One corrupt line does not erase the log.  Each line is
    decoded independently with \texttt{DisallowUnknownFields}
    (trust-boundary symmetric with reflection and result loaders);
    a failing line produces a warning while the rest of the file
    decodes.
  \item One oversized line does not truncate the tail.
    \texttt{bufio.Reader.ReadString} replaces \texttt{bufio.Scanner}
    so a line larger than any fixed cap no longer silently truncates
    every subsequent line; memory is bounded by the 16 MiB
    whole-file cap instead.
  \item One attacker-planted ESC sequence does not reach the
    operator terminal.  A \texttt{sanitizeWarning} helper walks
    warning strings rune-by-rune, escaping bytes \texttt{< 0x20}
    (except tab/space), DEL, and C1 (U+007F--U+009F) while
    preserving legitimate multi-byte UTF-8.  Sanitization runs at
    every warning append site in \texttt{decodeEventLog} --- the
    source --- so CLI, MCP, and the DES-020 formatter all consume
    clean strings without double-escaping.
  \item TOCTOU-resistant whole-file cap.  A single-fd \texttt{os.Open}
    $\to$ \texttt{f.Stat()} $\to$
    \texttt{io.ReadAll(io.LimitReader(f, maxLogSize+1))} sequence
    pins the inode so a concurrent writer cannot replace or
    redirect the file between stat and read; the \texttt{+1}
    overflow byte turns silent cap-bypass-via-race into a distinct
    ``grew past cap'' error.
\end{itemize}

\textbf{Trust boundary.}  Per-line strict decode
(\texttt{DisallowUnknownFields}), \field{ts} parseability check at
decode time, \texttt{missionIDPattern} validation at API entry,
directory rejection via \texttt{info.IsDir}, whole-file size cap at
16 MiB.  Every defense the Phase 3.4 / 3.6 loaders enforce, the
Phase 3.7 reader enforces --- and two the siblings don't (line-
length resilience and attacker-controlled-byte sanitization).

\textbf{Forward-compat for unknown event types.}  Event type strings
are preserved as opaque during decode.  A future phase
(\texttt{worker\_spawned}, \texttt{round\_started}, etc.) can emit
new types without a reader change.  The \texttt{--event <type,list>}
CLI filter accepts arbitrary strings and simply returns empty if no
events match --- it does not validate against a closed enum.

\subsection{Evaluator Discipline}

Three Phase 3 invariants together close the ``silent drift'' failure
modes the architecture documents named:

\begin{enumerate}
  \item \textbf{Evaluator content drift} (DES-033) --- the verifier
    content hash gate refuses spawns when the evaluator's identity
    files have changed since pinning.
  \item \textbf{Round drift} (DES-034) --- the round-advance gate
    refuses to begin round N+1 without a reflection for round N
    and refuses to exceed \field{budget.rounds}.
  \item \textbf{Worker bias} (DES-035) --- the role-overlap check
    refuses missions where the worker and evaluator share a role,
    and the verifier context isolator strips parent transcript and
    worker scratch from the verifier subagent.
\end{enumerate}

The mission close gate (DES-036) is the fourth, structural one:
no terminal transition without a typed result artifact.  The
event log reader (DES-037) is the post-mortem substrate that lets
an operator reconstruct what happened when any of the above fired.

% ===================================================================
\section{Agent Generation}
% ===================================================================

Generated agent definition files in \fpath{.claude/agents/<handle>.md}
are produced by \texttt{GenerateAgentFiles()} from ethos identity,
personality, writing style, role, and team data.  The result is a
standard Claude Code agent definition that another agent
(\texttt{Task} or \texttt{Agent} primitive) can spawn directly.

\subsection{Frontmatter}

YAML frontmatter carries:

\begin{itemize}
  \item \field{name} --- identity handle
  \item \field{description} --- the opening line of the persona
    block, suitable for an agent picker
  \item \field{tools} --- the role's \field{tools} field, transformed
    into the Claude Code agent tools allowlist
  \item \field{model} --- the role's \field{model} field
    (\texttt{opus}, \texttt{sonnet}, \texttt{haiku},
    \texttt{inherit}, or full \texttt{claude-*} ID); empty means
    inherit from the parent session
  \item \field{skills} --- always includes \texttt{baseline-ops}
    (Phase 2.4) so every generated agent gets the operational
    discipline subset (Read instead of cat, no commits without
    permission, scope discipline, security basics)
  \item \field{hooks} --- role-based PostToolUse hooks (Phase 2.2):
    implementation roles get \texttt{make check} after every
    \texttt{Write} or \texttt{Edit}; review-only roles get no
    hooks because tool restrictions already prevent writes
\end{itemize}

\subsection{Anti-Responsibility Derivation (Phase 2.1)}

The agent body includes a ``What You Don't Do'' section derived
mechanically from the team collaboration graph.  For each
\texttt{reports\_to} edge from the agent's role to a target role,
the target role's \field{responsibilities} list contributes one
bullet to the generated section, with parenthetical attribution to
the target.

Concretely, if \texttt{go-specialist} \texttt{reports\_to}
\texttt{coo} and \texttt{coo}'s responsibilities include
``execution quality and velocity across all engineering'', the
generated \texttt{bwk.md} contains:

\begin{verbatim}
## What You Don't Do

You report to coo. These are not yours:

- execution quality and velocity across all engineering (coo)
- sub-agent delegation and review (coo)
- release management (coo)
- operational decisions (coo)
\end{verbatim}

The preamble names the target(s) with Oxford comma joining for
three or more.  Bullet attribution is emitted on every bullet
including the $N{=}1$ case for consistency and grep-ability.

The derivation is mechanical --- no pattern matching, no LLM, no
verb-phrase transformation.  Responsibility strings are emitted
verbatim after whitespace normalization (trimmed edges, internal
newlines collapsed to spaces, empty-after-trim entries skipped
with a stderr warning).  The MVP handles only \texttt{reports\_to}
edges; \texttt{collaborates\_with} and \texttt{delegates\_to}
edges from the agent's role produce a stderr warning and no
bullets --- their semantics (peer, downward) deserve different
phrasing and are deferred.  Missing target roles are skipped with a
stderr warning at generation time; the rest of the file still
renders.

Anti-responsibilities prevent scope creep and enforce the
architecture rule that a role is defined by its inputs, outputs,
and \emph{what it does not touch}.  No schema change is required:
the \field{Role} struct has no \field{anti\_responsibilities}
field --- the generator reads the team graph and the target
roles' own \field{responsibilities} lists.

\subsection{Structured Output Format (Phase 2.3)}

Roles with an \field{output\_format} field include the template in
the agent body as an ``Output Format'' section.  Output formats
include structured fields (files changed, verdict, confidence,
open questions) alongside human-readable prose --- the same shape
the mission contract's \texttt{Result} artifact uses, so worker
output composes cleanly into a \cmd{ethos mission result}
submission.

\subsection{Mission Skill (Phase 2.5/2.6)}

The \cmd{/mission} skill (deployed at
\fpath{\~{}/.claude/skills/mission/SKILL.md}) scaffolds mission-
shaped delegations interactively:

\begin{enumerate}
  \item Resolve the agent from the team roster (show available
    agents, roles, status via the \tool{team} and \tool{identity}
    MCP tools).
  \item Build the mission contract: task, inputs, outputs, success
    criteria, write-set, constraints --- pre-populated from
    conversation context, presented as confirmable options.
  \item Check for write-set conflicts with running missions via the
    \tool{mission} MCP tool's \texttt{list} method (Phase 3.2).
  \item Spawn the agent via Claude Code's \texttt{Agent} primitive
    with the structured prompt that includes the mission ID.
  \item Optionally track the mission in beads.
\end{enumerate}

The skill sits at the structured layer between CLAUDE.md guidance
(``delegate to specialists'') and the raw \texttt{Agent} primitive,
turning mission discipline into a single command instead of an ad
hoc paragraph.

\subsection{Idempotence}

\texttt{GenerateAgentFiles} is idempotent: a generated file is only
written when its content differs from the existing file on disk.
This prevents unnecessary file churn and git noise on every
\texttt{SessionStart}.  The main agent and human identities are
skipped --- humans don't get \fpath{.claude/agents} entries, and
the main agent is whatever the user is currently running as.

% ===================================================================
\section{Operational Hooks}
% ===================================================================

Phase 4 deepens the hook system with three additions: working
context at \texttt{SessionStart}, role-based pre-tool safety
constraints, and \texttt{PostToolUse} audit logging.  These are
additive --- they layer onto the existing persona-animation
infrastructure without changing it.

\subsection{Working Context at SessionStart (Phase 4.1)}

\texttt{HandleSessionStart} emits a \texttt{\#\# Working Context}
block alongside the persona and team blocks.  Contents:

\begin{itemize}
  \item Current git branch and short commit
  \item Uncommitted changes (file count and basic shape)
  \item Recent issues from the bead store (open beads assigned to
    the resolved persona)
  \item Recent missions from \cmd{ethos mission list --status open}
\end{itemize}

The block is generated server-side in the hook and included in the
\texttt{additionalContext} payload for the SessionStart event.  No
network calls, no subprocess spawning beyond what
\texttt{git status --porcelain} already costs.

\subsection{Role-Based Pre-Tool Safety (Phase 4.2)}

Roles carry an optional \field{safety\_constraints} field --- a
list of pattern strings that the pre-tool hook reads.  A
\texttt{reviewer} role denies \texttt{Write}, \texttt{Edit}, and
\texttt{NotebookEdit}.  A \texttt{researcher} role denies
destructive \texttt{Bash} patterns (\texttt{rm -rf},
\texttt{git push --force}).  An \texttt{architect} role denies all
write tools.

The hook reads the active subagent's role from the session roster,
loads the role's \field{safety\_constraints}, and refuses the tool
invocation with a structured error.  Tool restrictions in the
generated agent's frontmatter (Phase 2.2) are the first line of
defense; the safety constraints field is the second, enforced even
when an agent is invoked without a generated definition file.

\subsection{PostToolUse Audit Logging (Phase 4.3)}

\texttt{HandleFormatOutput} (the existing \texttt{PostToolUse} hook
that produces DES-020 two-channel formatted output) gains an
optional audit log writer.  Every tool invocation appends a JSON
line to \fpath{\~{}/.punt-labs/ethos/sessions/<id>.audit.jsonl}:

\begin{verbatim}
{"ts":"2026-04-08T15:23:00Z","tool":"Write","persona":"bwk",
 "mission":"m-2026-04-08-005","args_summary":"path=internal/foo.go"}
\end{verbatim}

The audit log is per-session, append-only, and uses the same
\texttt{flock}-protected write pattern as the mission event log.
Mission attribution comes from the active session participant's
ext map.  Like the mission event log, sanitization runs at the
write site so no attacker-controlled byte reaches a downstream
consumer.

% ===================================================================
\section{Cross-Tool Integration}
% ===================================================================

Ethos is the identity anchor.  Other Punt Labs tools read ethos
state from the filesystem to configure themselves.  No tool imports
ethos as a Go library --- the file format is the contract.

\subsection{Vox: Voice Configuration via Extensions}

Vox stores voice configuration in
\fpath{<handle>.ext/vox.yaml}.  Keys include \field{default\_mood},
\field{voice\_id}, and \field{session\_context} (a behavioral
instruction injected into the model's context at SessionStart per
DES-022).  Vox reads its own namespace; ethos validates the
namespace structure but never interprets the values.

\subsection{Beadle: Identity-Aware Email}

Beadle reads \fpath{<handle>.ext/beadle.yaml} for IMAP credentials,
GPG key bindings, and signature blocks.  When an agent
impersonating the COO sends email, Beadle resolves the active
identity via \texttt{ethos whoami}, loads the COO's email and GPG
key from the identity's beadle ext namespace, and signs the message
with the right key.  The same identity that ships commits as
\texttt{Claude Agento} signs email as \texttt{claude@punt-labs.com}
--- one persona, multiple channels.

\subsection{Biff: Identity-Aware Team Messaging}

Biff reads \fpath{<handle>.ext/biff.yaml} for the agent's preferred
TTY and presence configuration.  Biff's \cmd{/who} and
\cmd{/finger} commands resolve identities through ethos so a
broadcast like \cmd{biff write @bwk ``rebase done''} routes to the
session where \texttt{bwk} is currently active.  Biff also reads
the team registry under \fpath{teams/} for collaboration graphs ---
the same data that drives anti-responsibility derivation in
generated agent files.

\subsection{Pattern: Identity-Aware Channel Binding}

The cross-tool pattern is \emph{identity-aware channel binding}:
the identity is durable, the channel binding is the verb (sends
email, signs commits, broadcasts a message, speaks aloud), and
ethos owns the binding's storage.  When the same agent shows up in
five tools, the user sees one persona --- the same name, the same
voice, the same writing style, the same email address --- because
each tool reads the same identity record.  This is the operational
half of the ``AI agents as first-class engineering citizens''
vision.

% ===================================================================
\section{Storage Layout}
% ===================================================================

\begin{center}
\begin{longtable}{@{}lll@{}}
\toprule
\textbf{Scope} & \textbf{Path} & \textbf{Git-tracked?} \\
\midrule
\endfirsthead
\toprule
\textbf{Scope} & \textbf{Path} & \textbf{Git-tracked?} \\
\midrule
\endhead
\multicolumn{3}{@{}l}{\emph{Repo-local (under repository root):}} \\
\midrule
Repo identities  & \fpath{.punt-labs/ethos/identities/<handle>.yaml}
                 & Yes \\
Repo talents     & \fpath{.punt-labs/ethos/talents/<slug>.md}
                 & Yes \\
Repo personalities & \fpath{.punt-labs/ethos/personalities/<slug>.md}
                 & Yes \\
Repo writing styles & \fpath{.punt-labs/ethos/writing-styles/<slug>.md}
                 & Yes \\
Repo roles       & \fpath{.punt-labs/ethos/roles/<slug>.yaml}
                 & Yes \\
Repo teams       & \fpath{.punt-labs/ethos/teams/<slug>.yaml}
                 & Yes \\
Repo agents      & \fpath{.punt-labs/ethos/agents/<name>.md}
                 & Yes \\
Repo config      & \fpath{.punt-labs/ethos.yaml}
                 & Yes \\
\midrule
\multicolumn{3}{@{}l}{\emph{User-global
  (\fpath{\~{}/.punt-labs/ethos/}):}} \\
\midrule
Global identities  & \fpath{\~{}/.punt-labs/ethos/identities/<handle>.yaml}
                   & No \\
Extensions         & \fpath{\~{}/.punt-labs/ethos/identities/<handle>.ext/<ns>.yaml}
                   & No \\
Global talents     & \fpath{\~{}/.punt-labs/ethos/talents/<slug>.md}
                   & No \\
Global personalities & \fpath{\~{}/.punt-labs/ethos/personalities/<slug>.md}
                   & No \\
Global writing styles & \fpath{\~{}/.punt-labs/ethos/writing-styles/<slug>.md}
                   & No \\
Global roles       & \fpath{\~{}/.punt-labs/ethos/roles/<slug>.yaml}
                   & No \\
Global teams       & \fpath{\~{}/.punt-labs/ethos/teams/<slug>.yaml}
                   & No \\
\midrule
\multicolumn{3}{@{}l}{\emph{Sessions and missions
  (operational state):}} \\
\midrule
Sessions           & \fpath{\~{}/.punt-labs/ethos/sessions/<id>.yaml}
                   & No \\
Session locks      & \fpath{\~{}/.punt-labs/ethos/sessions/<id>.lock}
                   & No \\
Current session    & \fpath{\~{}/.punt-labs/ethos/sessions/current/<pid>}
                   & No \\
Session audit log  & \fpath{\~{}/.punt-labs/ethos/sessions/<id>.audit.jsonl}
                   & No \\
Mission contracts  & \fpath{\~{}/.punt-labs/ethos/missions/<id>.yaml}
                   & No \\
Mission event logs & \fpath{\~{}/.punt-labs/ethos/missions/<id>.jsonl}
                   & No \\
Mission reflections & \fpath{\~{}/.punt-labs/ethos/missions/<id>.reflections.yaml}
                   & No \\
Mission results    & \fpath{\~{}/.punt-labs/ethos/missions/<id>.results.yaml}
                   & No \\
Mission locks      & \fpath{\~{}/.punt-labs/ethos/missions/<id>.lock}
                   & No \\
Mission counter    & \fpath{\~{}/.punt-labs/ethos/missions/.counter-<date>}
                   & No \\
Mission create lock & \fpath{\~{}/.punt-labs/ethos/missions/.create.lock}
                   & No \\
Generated agents   & \fpath{.claude/agents/<handle>.md}
                   & No \\
\bottomrule
\end{longtable}
\end{center}

All user-global state lives under \fpath{\~{}/.punt-labs/ethos/}.
Repo-scoped state lives under \fpath{.punt-labs/ethos/} in the
repository root (typically a git submodule of the team registry).
Mission state is operational (per-machine, short-lived, not
configuration) and lives under the global root.

% ===================================================================
\section{CLI Surface}
% ===================================================================

The binary exposes subcommands grouped by concern.  All commands
accept a global \cmd{--json} flag for machine-readable output.

\subsection{Identity Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{whoami}              & Show the caller's identity (resolved
                            via 4-step chain). \\
\cmd{identity}            & Parent command (defaults to whoami). \\
\cmd{identity whoami}     & Show the caller's identity. \\
\cmd{identity list}       & List all identities from both stores. \\
\cmd{identity get <h>}    & Show full identity details including
                            extensions and resolved content. \\
\cmd{identity create}     & Create identity interactively or from
                            a YAML file (\cmd{-f}). \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Attribute Commands}

Three parallel command groups: \cmd{talent}, \cmd{personality},
\cmd{writing-style}.  Each provides:

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Subcommand} & \textbf{Description} \\
\midrule
\cmd{create <slug>}       & Create from \$EDITOR or file (\cmd{-f}). \\
\cmd{list}                & List all slugs from both stores. \\
\cmd{show <slug>}         & Print markdown content. \\
\cmd{delete <slug>}       & Remove the \texttt{.md} file. \\
\cmd{set <handle> <slug>} & Bind to an identity (personality, writing
                            style only). \\
\cmd{add <handle> <slug>} & Add to identity (talents only). \\
\cmd{remove <h> <slug>}   & Remove from identity (talents only). \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Role Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{role create <name>}  & Create interactively or from YAML
                            (\cmd{-f}). \\
\cmd{role list}            & List all roles from both stores. \\
\cmd{role show <name>}     & Show role details. \\
\cmd{role delete <name>}   & Delete (guarded by referential
                            integrity). \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Team Commands}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{team create <name>}                & Create from YAML (\cmd{-f})
                                          or interactive. \\
\cmd{team list}                          & List all teams. \\
\cmd{team show <name>}                   & Show team details. \\
\cmd{team delete <name>}                 & Delete a team. \\
\cmd{team add-member <t> <id> <role>}    & Add member with referential
                                          integrity check. \\
\cmd{team remove-member <t> <id> <role>} & Remove a member. \\
\cmd{team add-collab <t> <f> <to> <type>} & Add directed
                                          collaboration. \\
\cmd{team for-repo [repo]}               & Show teams for a repo
                                          (defaults to current). \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Mission Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{mission create -f <yaml>}      & Create a mission contract.
                                       Strict YAML decode; runs
                                       write-set conflict scan,
                                       role-overlap check, and
                                       evaluator hash computation. \\
\cmd{mission show <id-or-prefix>}   & Show contract, current round,
                                       reflection log, and result
                                       log.  \cmd{--json} emits
                                       \texttt{ShowPayload}. \\
\cmd{mission list}                  & List missions.  \cmd{--status
                                       open$|$closed$|$failed$|$escalated$|$all}
                                       (default: open). \\
\cmd{mission close <id>}            & Transition to terminal status.
                                       Refused without a valid
                                       result artifact for the
                                       current round. \\
\cmd{mission reflect <id> -f <yaml>} & Submit a typed reflection
                                       for the current round.
                                       Append-only. \\
\cmd{mission advance <id>}          & Bump current round.  Refused
                                       without a non-terminal
                                       reflection for the current
                                       round, and refused on budget
                                       exhaustion. \\
\cmd{mission reflections <id>}      & Print the round-by-round
                                       reflection log. \\
\cmd{mission result <id> -f <yaml>}  & Submit a typed result for the
                                       current round.  Append-only.
                                       Validates files\_changed
                                       against the contract
                                       write-set. \\
\cmd{mission results <id>}          & Print the round-by-round
                                       result log. \\
\cmd{mission log <id>}              & Print the JSONL event log
                                       in on-disk order.
                                       \cmd{--event <type,list>},
                                       \cmd{--since <RFC3339>}.
                                       Warnings appended on
                                       partial decode. \\
\bottomrule
\end{tabular}
\end{center}

The mission subcommands form the operational interface to Phase 3
discipline.  Read-only subcommands (\cmd{show}, \cmd{list},
\cmd{reflections}, \cmd{results}, \cmd{log}) use the bare
\texttt{missionStore} helper.  \cmd{mission create} uses
\texttt{missionStoreForCreate}, which constructs a live
\texttt{RoleLister} and \texttt{HashSources} so the role-overlap
check (DES-035) and the evaluator content hash (DES-033) fire on
the production path.

\subsection{Extension Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{ext get <p> <ns> [key]}       & Read one key or all keys in a
                                     namespace. \\
\cmd{ext set <p> <ns> <key> <val>} & Write a key-value pair. \\
\cmd{ext del <p> <ns> [key]}       & Delete one key or an entire
                                     namespace. \\
\cmd{ext list <p>}                 & List all namespaces for a
                                     persona. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Session Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{iam <persona>}            & Declare persona in current session
                                 (auto-detects session ID and agent
                                 ID). \\
\cmd{session}                  & Show current session roster. \\
\cmd{session create}           & Create a new roster (called by
                                 SessionStart hook). \\
\cmd{session join}             & Add a participant (called by
                                 SubagentStart hook). \\
\cmd{session leave}            & Remove a participant (called by
                                 SubagentStop hook). \\
\cmd{session write-current}    & Write PID-keyed session file. \\
\cmd{session delete-current}   & Delete PID-keyed session file. \\
\cmd{session purge}            & Clean up stale rosters. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Hook Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{hook session-start}    & SessionStart hook handler. \\
\cmd{hook session-end}      & SessionEnd hook handler. \\
\cmd{hook subagent-start}   & SubagentStart hook handler.
                              Includes the DES-033 verifier hash
                              gate and the DES-035 verifier
                              context isolator. \\
\cmd{hook subagent-stop}    & SubagentStop hook handler. \\
\cmd{hook pre-compact}      & PreCompact hook handler. \\
\cmd{hook format-output}    & PostToolUse output formatter and
                              (Phase 4.3) audit log writer. \\
\bottomrule
\end{tabular}
\end{center}

The \cmd{hook} group is hidden from help output.  These commands are
invoked by Claude Code hook configuration, not by users.

\subsection{Admin Commands}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\cmd{version}              & Print version string. \\
\cmd{resolve-agent}        & Print the default agent handle from repo
                             config. \\
\cmd{doctor}               & Check installation health. \\
\cmd{serve}                & Start MCP server (stdio transport). \\
\cmd{completion <shell>}   & Generate shell completion script. \\
\cmd{uninstall [--purge]}  & Remove plugin; \cmd{--purge} also
                             removes binary and data. \\
\bottomrule
\end{tabular}
\end{center}

% ===================================================================
\section{MCP Tool Surface}
% ===================================================================

\texttt{ethos serve} starts a stdio-transport MCP server that
registers up to 10~tools via \texttt{Handler.RegisterTools()}.
The role, team, and mission tools are registered when their stores
are configured; in a full deployment all 10 are present.  Each tool
is a consolidated endpoint that accepts a \field{method} parameter
for dispatch.  The \texttt{Handler} struct holds injected stores:
\texttt{identity.IdentityStore}, \texttt{session.Store},
\texttt{attribute.Store} (talents, personalities, writing styles),
\texttt{role.LayeredStore}, \texttt{team.LayeredStore}, and
\texttt{mission.Store}.

\subsection{Identity Tool}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{whoami}  & Resolve caller's identity via 4-step chain. \\
\texttt{list}    & List all identities from both stores. \\
\texttt{get}     & Load full identity by handle. \\
\texttt{create}  & Create identity from provided fields. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Extension Tool}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{get}   & Read key(s) in a namespace. \\
\texttt{set}   & Write a key-value pair. \\
\texttt{del}   & Delete a key or namespace. \\
\texttt{list}  & List namespaces for a persona. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Session Tool}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{roster} & Load full roster (session ID auto-discovered if
                  omitted). \\
\texttt{iam}    & Declare persona for a participant. \\
\texttt{join}   & Register a participant. \\
\texttt{leave}  & Deregister a participant. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Attribute Tools}

Three tools --- \tool{talent}, \tool{personality},
\tool{writing\_style} --- each with \texttt{create}, \texttt{list},
\texttt{show}, \texttt{delete}, \texttt{set}, \texttt{add},
\texttt{remove}.

\subsection{Role Tool}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{create}  & Create a new role. \\
\texttt{list}    & List all role names. \\
\texttt{show}    & Load role details. \\
\texttt{delete}  & Delete role (referential integrity guarded). \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Team Tool}

\begin{center}
\begin{tabular}{@{}lp{6cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{create}         & Create a new team with members. \\
\texttt{list}           & List all team names. \\
\texttt{show}           & Load full team definition. \\
\texttt{delete}         & Delete a team. \\
\texttt{add\_member}    & Add identity/role pair. \\
\texttt{remove\_member} & Remove identity/role pair. \\
\texttt{add\_collab}    & Add directed collaboration. \\
\texttt{for\_repo}      & Find teams that own a repository. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Mission Tool}

\begin{center}
\begin{tabular}{@{}lp{6.5cm}@{}}
\toprule
\textbf{Method} & \textbf{Description} \\
\midrule
\texttt{create}      & Create a contract from a YAML payload.
                       Symmetric with the CLI \cmd{create}: same
                       strict decode, same write-set conflict scan,
                       same role-overlap check, same hash
                       computation. \\
\texttt{show}        & Load contract, results, and warnings as a
                       \texttt{ShowPayload}. \\
\texttt{list}        & List missions filtered by status. \\
\texttt{close}       & Terminal transition.  Refused without a
                       valid result artifact for the current
                       round. \\
\texttt{reflect}     & Append a typed reflection. \\
\texttt{reflections} & Read the round-by-round reflection log. \\
\texttt{advance}     & Bump current round.  Same gate as the CLI
                       \cmd{advance}. \\
\texttt{result}      & Append a typed result.  Same validation as
                       the CLI \cmd{result}. \\
\texttt{results}     & Read the round-by-round result log. \\
\texttt{log}         & Read the JSONL event log.  \field{event} and
                       \field{since} filters; wrapped JSON envelope
                       \texttt{\{events: [...], warnings: [...]\}}. \\
\bottomrule
\end{tabular}
\end{center}

The MCP \tool{mission} tool is the symmetric counterpart of the CLI
\cmd{mission} subcommands.  Every gate fires identically regardless
of which surface called the store: an MCP caller cannot bypass the
write-set conflict scan, the evaluator hash, or the close gate.

\subsection{Doctor Tool}

Standalone admin tool with no methods.  Runs all health checks
(identity directory, human identity, default agent, duplicate
fields) and returns a formatted status table.

\subsection{Output Formatting}

The \texttt{FormatOutput} hook (DES-020) provides two-channel
display for every MCP tool response: a compact summary in
\texttt{updatedMCPToolOutput} and full data in
\texttt{additionalContext}.  Each tool has a dedicated formatter
in \fpath{internal/hook/format\_output.go} that produces columnar
tables (identity list, session roster, role/team details, mission
list, reflection log, result log, event log) or structured
summaries.  New tools must ship with a formatter before the MCP
tool is registered --- DES-020 makes this a hard rule.

% ===================================================================
\section{Design Invariants}
% ===================================================================

These invariants are enforced by the architecture and hold across
all operations:

\begin{enumerate}
  \item \textbf{Sidecar contract.}  No consumer imports ethos.
    The contract is file paths and YAML formats.

  \item \textbf{Unified schema.}  Humans and agents use the same
    \texttt{Identity} struct.  The \field{kind} field is the only
    difference.  No separate code paths.

  \item \textbf{No consumer-specific fields.}  Ethos will never add
    \field{gpg\_key\_id}, \field{preferred\_tty}, or any field that
    serves a single consumer.  The extension system exists for this.

  \item \textbf{Atomic identity creation.}  \texttt{Store.Save}
    opens the file with \texttt{O\_EXCL}.  Two concurrent creates
    of the same handle: one succeeds, one fails.  No partial writes.

  \item \textbf{Atomic roster writes.}  Roster files are written to
    a temp file, then renamed.  No reader sees a partial roster.

  \item \textbf{Flock concurrency.}  All roster, identity, and
    mission mutations acquire \texttt{flock(LOCK\_EX)} on a
    per-target lock file.  No daemon, no database, no race
    conditions on concurrent hooks or concurrent CLI invocations.

  \item \textbf{Extension isolation.}  Each consumer's extension
    data lives in its own YAML file.  Concurrent writes by different
    tools cannot corrupt each other's data.

  \item \textbf{Process tree caching.}  \texttt{FindClaudePID()}
    caches the result for the process lifetime via
    \texttt{sync.Once}.

  \item \textbf{No shell execution.}  All subprocess invocations use
    argument lists, never shell interpretation.

  \item \textbf{Depth-bounded tree walk.}  The process tree walk
    caps at 10~iterations.

  \item \textbf{Layered resolution.}  Repo-local takes priority for
    identity, attribute, role, and team lookup.  Extensions are
    always global.  Mission state is always global (operational,
    not configuration).

  \item \textbf{Referential integrity.}  Teams enforce that every
    member's identity and role exist at creation and mutation time.
    Roles cannot be deleted while referenced by any team.
    Collaborations must reference roles filled by current members.

  \item \textbf{Idempotent agent generation.}  Generated agent files
    are only written when their content differs from the existing
    file on disk.

  \item \textbf{Mission contract totality} (DES-031).  Every
    contract has a non-empty write-set, a non-empty success
    criteria list, an evaluator handle pinned with a server-
    controlled timestamp, and a budget in $[1, 10]$.  Every
    mission ID matches \verb|^m-\d{4}-\d{2}-\d{2}-\d{3}$|.

  \item \textbf{Append-only event log} (DES-031, DES-037).  The
    JSONL event log is the source of truth for what happened to a
    mission.  Every state transition lands as one line under the
    per-mission flock with \texttt{O\_APPEND}.  The writer is
    private and frozen; the reader degrades line-by-line and
    sanitizes warning strings at source.

  \item \textbf{Trust-boundary symmetry} (DES-031, DES-036, DES-037).
    Every gate that fires on the write path also fires on the read
    path.  \texttt{KnownFields(true)} on every YAML decode;
    \texttt{DisallowUnknownFields} on every JSONL decode;
    \texttt{missionIDPattern} validation at API entry.

  \item \textbf{Frozen evaluator content hash} (DES-033).  The
    evaluator's personality, writing style, talents, and team-scoped
    role bindings are content-hashed at mission create time and
    re-hashed at every verifier subagent spawn.  Any drift refuses
    the spawn.

  \item \textbf{Bounded rounds} (DES-034).  \field{Budget.Rounds}
    is a hard cap.  No round N+1 begins without a non-terminal
    reflection for round N.  No silent seventh round.

  \item \textbf{Verifier independence} (DES-035).  Worker and
    evaluator handles cannot share a team-scoped role binding.
    Verifier subagents receive only the mission contract,
    verification criteria, and file allowlist --- the persona block,
    parent transcript, and worker scratch are excluded.

  \item \textbf{Close gate totality} (DES-036).  No mission
    transitions to a terminal status without a valid result
    artifact for the current round.  No override flag.
\end{enumerate}

% ===================================================================
\section{Security Boundaries}
% ===================================================================

\subsection{Filesystem Boundary}

\begin{itemize}
  \item \textbf{User-global state}: \fpath{\~{}/.punt-labs/ethos/} ---
    identities, extensions, sessions, missions.  Standard user-mode
    file permissions.
  \item \textbf{Repo-local state}: \fpath{.punt-labs/ethos/} ---
    identities, attributes, roles, teams, agents.  Git-tracked;
    visible to collaborators.
  \item \textbf{No secrets.}  Ethos stores no passwords, API keys,
    or tokens.  Sensitive bindings (GPG keys, IMAP passwords) live
    in consumer extensions or credential stores, not in ethos.
\end{itemize}

\subsection{Process Boundary}

\begin{itemize}
  \item \textbf{MCP}: stdio transport.  No network listener.
  \item \textbf{Native proc reading}: \texttt{/proc/<pid>/stat} on
    Linux, \texttt{sysctl kern.proc.pid} on macOS.  Read-only.
    Signal~0 for process liveness checks (no side effects).
\end{itemize}

\subsection{Trust Boundary}

\begin{itemize}
  \item \textbf{Identity data is self-declared.}  Ethos does not
    verify that the email, GitHub handle, or voice binding is
    authentic.  Verification is the responsibility of consumers
    (Beadle verifies PGP, Biff verifies GitHub tokens).
  \item \textbf{Extension data is opaque.}  Ethos enforces
    structural constraints (namespace format, key format, size
    limits) but does not validate or interpret values.
  \item \textbf{Path traversal prevention.}  Attribute slugs and
    mission IDs are validated against strict patterns and resolved
    paths are checked to remain within the target directory.
    Handle-based file paths use \texttt{filepath.Base} to prevent
    directory escape.
\end{itemize}

\subsection{Phase 3 Trust Boundaries}

The mission primitive adds a small number of new trust boundaries.
Each is enforced symmetrically on read and write paths.

\begin{itemize}
  \item \textbf{Write-set admission} (DES-032).  A new mission whose
    \field{write\_set} segment-prefix-overlaps any open mission's
    \field{write\_set} is rejected at \texttt{Store.Create}.  The
    overlap normalization filters empty segments
    (\texttt{internal//foo}) and single-dot segments
    (\texttt{./internal/foo}) so the equivalence class is closed.
  \item \textbf{Strict on-disk YAML decode} (DES-031).
    \texttt{Store.Load} and \texttt{Store.loadLocked} call
    \texttt{Validate()} for defense in depth on reads.  Both the
    CLI and MCP create paths and the load path use
    \texttt{KnownFields(true)} so an attacker with local write
    access cannot smuggle extra fields.
  \item \textbf{Control-byte sanitization} (DES-031, DES-037).
    Identity handles, slugs, and mission IDs reject ASCII control
    bytes.  Mission write-set entries reject \texttt{..}, absolute
    paths, drive letters, UNC, null bytes, and C0 control
    characters.  The event log reader sanitizes warning strings at
    source so no attacker-controlled byte reaches operator
    terminals or MCP consumers.
  \item \textbf{Mission ID pattern validation} (DES-037).  Every
    mission loader validates the mission ID against
    \verb|^m-\d{4}-\d{2}-\d{2}-\d{3}$| at API entry, before any
    \texttt{filepath.Base} collapse, so raw attacker-controlled
    bytes never reach a downstream \texttt{*fs.PathError} string.
  \item \textbf{Symmetric file-trust on results and events} (DES-036
    round 5, DES-037).  Both \texttt{LoadResults} and
    \texttt{LoadEvents} cross-check the on-disk mission identity
    against the requested ID --- the same trust-boundary symmetry
    DES-031 round 3 introduced for the contract itself.
  \item \textbf{Symlink follow weakness (filed as
    \texttt{ethos-jjm}).}  All four mission loaders
    (\texttt{LoadEvents}, \texttt{LoadReflections},
    \texttt{LoadResults}, \texttt{Store.Load}) currently follow
    symlinks via \texttt{os.ReadFile} or \texttt{os.Open} without
    checking whether the resolved target is inside the missions
    directory.  Fixed in a coordinated change across all four
    loaders.
\end{itemize}

% ===================================================================
\section{Key Properties}
% ===================================================================

The architecture guarantees these properties, verifiable by
inspection and testing:

\begin{enumerate}
  \item \textbf{Resolution totality}: \texttt{Resolve()} always
    returns a handle or an error.  No silent empty-string fallback.

  \item \textbf{Extension namespace isolation}: concurrent writes by
    different tools target different files.

  \item \textbf{Session liveness}: \texttt{Purge()} uses process
    signal~0 for numeric PIDs and age-based TTL for non-numeric IDs.

  \item \textbf{Hook-driven lifecycle}: the session roster is fully
    managed by Claude Code hook events.  No polling, no daemon, no
    manual intervention.

  \item \textbf{Cold start performance}: Go binary with no runtime
    dependencies.  $\sim$10\,ms cold start measured on Apple~M2.

  \item \textbf{Store injectability}: all CRUD operations accept a
    store with a configurable root path.  Tests use
    \texttt{t.TempDir()}.  No hardcoded \texttt{\$HOME}.

  \item \textbf{Persona survivability}: \texttt{PreCompact} re-injects
    the full persona block, extension context, and team context.

  \item \textbf{Team integrity}: Z-spec invariants are enforced on
    every team write operation.  A team is never persisted in an
    invalid state.

  \item \textbf{Mission contract totality} (DES-031): every contract
    that survives \texttt{Validate()} has a non-empty write-set, a
    non-empty success-criteria list, a pinned evaluator, and a
    budget in $[1, 10]$.

  \item \textbf{Event log append-only guarantee} (DES-031, DES-037):
    every state transition appends one line under the per-mission
    flock.  Concurrent writers cannot interleave because the lock
    serializes them; the writer is private and unexported so no
    external caller can deadlock by appending from inside another
    locked block.

  \item \textbf{Close gate totality} (DES-036): no mission
    transitions to a terminal status without a valid result
    artifact for the current round.  This holds for CLI and MCP
    callers identically because the gate lives at the store
    boundary.

  \item \textbf{Evaluator reproducibility} (DES-033): two verifier
    spawns of the same evaluator handle against the same identity
    files compute the same content hash.  Hash inputs use byte-
    length prefixes (not rune counts) and ASCII control-byte
    separators that ethos identity validation rejects from handles
    and slugs, so no field-boundary collision is reachable from
    legal identity content.

  \item \textbf{Round budget enforceability} (DES-034): no round
    N+1 begins without a non-terminal reflection for round N.  The
    advance gate is the single point of enforcement; an operator
    cannot bypass it by editing the contract on disk because
    \texttt{Validate()} re-fires on every load.

  \item \textbf{Verifier independence} (DES-035): a worker handle
    and an evaluator handle that share a team-scoped role under
    canonicalization cannot be paired in a single mission.  The
    verifier subagent receives no parent transcript and no worker
    scratch; the persona block is replaced by a structured
    isolation block.

  \item \textbf{Post-mortem readability} (DES-037): the JSONL
    event log reader returns every parseable line plus a
    line-numbered warnings list, even when the file is partially
    damaged or contains attacker-planted control sequences.

  \item \textbf{In-process handler testability} (DES-042): every
    CLI handler uses \texttt{RunE} and returns errors instead of
    calling \texttt{os.Exit}.  Tests call handlers directly via
    \texttt{execHandler}, capture stdout/stderr, and assert on
    both output and error values.  This moved \pkg{cmd/ethos/}
    from 22\% to 64\% measurable coverage.

  \item \textbf{Behavioral test isolation} (DES-043): L4 tests
    spawn real Claude Code agents in fixture repos with isolated
    \texttt{HOME} directories, their own identity stores, and
    per-test mission contracts.  No shared state leaks between
    tests; no test reads the developer's global identity store.

  \item \textbf{Layered assertion strategy} (DES-043):
    deterministic assertions (event log structure, git diff,
    result YAML schema) catch protocol violations without LLM
    calls.  LLM-judged assertions catch persona constraint
    violations that require judgment.  Adversarial scenarios
    combine both to prove the system holds under pressure.
    Each layer is independently runnable.
\end{enumerate}

% ===================================================================
\section{Testability Architecture}
\label{sec:testability}
% ===================================================================

Ethos has 1,055 tests across four pyramid levels, a 1.54:1 test-to-production
line ratio, and 75.6\% total coverage.  This section describes the
infrastructure that makes those numbers possible.

\subsection{Test Pyramid}

The test pyramid has four levels.  Each level trades speed for fidelity.

\begin{center}
\begin{longtable}{@{}llrp{6cm}@{}}
\toprule
\textbf{Level} & \textbf{Name} & \textbf{Time} & \textbf{What it proves} \\
\midrule
\endfirsthead
\toprule
\textbf{Level} & \textbf{Name} & \textbf{Time} & \textbf{What it proves} \\
\midrule
\endhead
L1 & Content validation  & $<$1\,s   & Identity YAML, attribute slugs,
                                        and extension schemas are
                                        well-formed.  Runs as
                                        \cmd{validate-content} in CI. \\
L2 & CLI subprocess      & $\sim$2\,s & Each CLI command exits with the
                                        correct code and produces the
                                        expected stdout/stderr.  Spawns
                                        a real \cmd{ethos} binary. \\
L2.5 & In-process handler & $\sim$3\,s & Handler logic, gate
                                          enforcement, and error paths
                                          exercised via \texttt{RunE}
                                          without spawning a subprocess.
                                          DES-042. \\
L3 & MCP integration     & $\sim$5\,s & Every MCP tool exercised
                                        through a real stdio client
                                        connection.  Verifies the
                                        JSON-RPC dispatch, input
                                        validation, and output
                                        formatting. \\
L4 & Behavioral          & $\sim$2\,min & Real Claude Code agents
                                          execute against mission
                                          contracts.  Protocol
                                          compliance, persona
                                          adherence, and adversarial
                                          resistance.  DES-043. \\
\bottomrule
\end{longtable}
\end{center}

L1--L3 run on every commit via \cmd{make check}.  L4 runs daily in
CI via \fpath{.github/workflows/behavioral.yml} and on-demand via
\cmd{make test-behavioral} (requires \texttt{ANTHROPIC\_API\_KEY} and
\texttt{claude} CLI).

\subsection{In-Process Handler Testing (DES-042)}

Before DES-042, every CLI handler used \texttt{Run} (no error return)
and called \texttt{os.Exit} on failure.  This made handlers untestable
in-process: coverage tools cannot measure code that kills the process.

The refactor converts every handler to \texttt{RunE}.  Handlers return
errors; the single \texttt{main()} function prints errors with the
\texttt{ethos:} prefix and routes to exit code 1 (runtime) or 2
(usage).  A \texttt{silentError} type signals ``already reported ---
exit non-zero without printing,'' used by handlers like
\texttt{runDoctor} that format their own failure output.

Tests call handlers via \texttt{execHandler}, which builds a
\texttt{cobra.Command}, injects flags, redirects stdout/stderr to
buffers, and returns the error.  No subprocess spawn, no binary
build, no process overhead.  This moved \pkg{cmd/ethos/} from 22\% to
64\% measurable coverage.

\subsection{Behavioral Test Infrastructure (DES-043)}

The L4 behavioral tests live in \fpath{tests/behavioral/} behind a
\texttt{//go:build behavioral} tag, excluded from \cmd{make check}.

\paragraph{Harness.}
\texttt{SetupFixture} creates an isolated environment per test: a
\texttt{t.TempDir()} root containing a fake \texttt{HOME}, a
git-initialized Go repo with a counter package, global and repo-local
identity stores, personality/writing-style/talent files for the
evaluator, a team with worker and evaluator roles, and ethos MCP
config for \cmd{claude --bare}.  The \texttt{SpawnAgent} function
runs \cmd{claude} as a subprocess with the fixture's environment,
a budget cap (\texttt{--max-budget-usd}), and a timeout.

\paragraph{Three Assertion Layers.}
Layer~A (deterministic) checks mission event logs, git diffs, and
result YAML structure --- no LLM calls required.  Layer~B
(LLM-judged) sends the agent's output and persona definition to
Claude Sonnet as a judge, which returns
\texttt{\{violated, evidence, confidence\}}.  Layer~C (adversarial)
deliberately tempts agents to break constraints, then combines
deterministic and judge assertions.  A low-confidence violation
($<0.7$) is logged as a warning, not a test failure.

\paragraph{Judge.}
\texttt{JudgePersonaCompliance} is a single \texttt{net/http} POST
to the Anthropic API.  No SDK dependency.  The prompt includes the
persona definition, the task, and the agent output; the response is
a structured JSON verdict.  The judge runs Claude Sonnet at
\textasciitilde\$0.05 per call.

\subsection{Convention Verification}

Convention tests enforce structural properties that code review
misses at scale:

\begin{itemize}
  \item Every CLI handler uses \texttt{RunE}, not \texttt{Run}.
  \item Every MCP tool has a corresponding formatter in
    \texttt{format\_output.go}.
  \item Identity YAML files pass strict decode
    (\texttt{KnownFields(true)}).
  \item Extension namespaces conform to the 32-char, lowercase
    pattern.
  \item Attribute slugs match the handle regex.
\end{itemize}

These are standard \texttt{go test} assertions that run in L1--L2.5.
When a new tool or handler is added, the convention test fails until
the corresponding formatter or \texttt{RunE} conversion is present.

\subsection{CI Integration}

Two workflows cover the pyramid:

\begin{enumerate}
  \item \textbf{Per-commit} (\fpath{.github/workflows/test.yml}):
    \cmd{make check} runs L1--L3.  Go vet, staticcheck, shellcheck,
    markdownlint, \texttt{go test -race -count=1 ./...}, and
    \cmd{validate-content}.
    Approximately 10 seconds wall time.
  \item \textbf{Daily behavioral}
    (\fpath{.github/workflows/behavioral.yml}):
    \cmd{make test-behavioral} runs L4.  Requires
    \texttt{ANTHROPIC\_API\_KEY} as a repository secret.  Each test
    spawns a Claude Code agent (\textasciitilde\$0.50,
    \textasciitilde 2\,min) so the suite is too slow and expensive
    for per-commit CI.
\end{enumerate}

The per-commit gate is the merge blocker.  The behavioral suite is
a confidence signal; failures create beads for investigation, not
automatic reverts.

\end{document}
